
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, cache = TRUE)
library(tidyverse)
library(ggrepel)
library(cowplot)
library(gridExtra)
library(spocc)
library(sf)
library(rnpn)
# library(devtools)
# devtools::install_github("bevingtona/planetR")
unloadNamespace("RJSONIO") # conflict with jsonlite
library(planetR)
library(httr)
library(jsonlite)
library(raster)
library(stringr)
library(pracma)
library(foreach)
library(doSNOW)
library(ptw)
library(EnvCpt)
library(rgdal)
library(mclust)
library(npreg)
library(nlme)
library(readxl)
library(fuzzyjoin)
library(taxize)
library(ptw)
library(segmented)
library(RhpcBLASctl)
library(geosphere)
library(nlme)
library(shiny)
library(rsconnect)
library(ggpubr)
library(scales)
```

* Taxa of interest are based on [Description and allergenic potential of 11 most important pollen taxa in the CUSSC region ranked by percent abundance relative to the sum of all pollen taxa over 31 NAB stations that meet inclusion criteria, 2003–2017] (https://link.springer.com/article/10.1007/s10453-019-09601-2/tables/2)
* Spring- and fall-flowering elms are treated separately as suggested by Allison and Yingxiao.
* I decide to separate grasses into early and late-flowering groups as well because many cities have two pollen peaks.
```{r, eval=TRUE}
taxa_list <- c("Quercus", "Cupressaceae", "Ambrosia", "Morus", "Pinaceae", "Ulmus early", "Ulmus late", "Fraxinus", "Betula", "Poaceae early", "Poaceae late", "Acer", "Populus")
taxa_short_list <- str_split(taxa_list, pattern = " ", simplify = T) [, 1]
```

* Ten major cities in CONUS with pollen count data and street tree inventory.
* Detroit only have location info of oak trees, grasses, and ragweeds.
```{r, eval=TRUE}
site_list <- c("NY", "SJ", "AT", "ST", "HT", "TP", "DT", "DV"#, "KC", "SL"
               )
sitename_list <- c("New York", "San Jose", "Austin", "Seattle", "Houston", "Tampa", "Detroit", "Denver"#, "Kansas City", "St. Louis"
                   )
```

* Focus on the life cycles from 2018 to 2021, spanning 2017 and 2022.
```{r, eval=TRUE}
year_list <- 2018:2021
```

# Data
## NAB data
### Read and clean data
```{r, eval=FALSE}
nab_path <- "/data/ZHULAB/phenology/nab/2021-10-04/"

# read in station coordinates
station_df <- read_csv("/data/ZHULAB/phenology/nab/NAB stations.csv") %>% # this csv is manually typed
  mutate(id = row_number())

file_list <- list.files(path = nab_path, pattern = ".xlsx", full.names = T)
nab_df_list <- vector(mode = "list", length = length(file_list))
for (i in 1:length(file_list)) {
  file <- file_list[i]

  # read in all data
  dat <- read_excel(
    file,
    col_names = F
  )
  start_n <- which(dat[, 1] == "Date") # excel files have headings of different number of rows. Search for the row(s) that starts with "Date" as the start of data table(s).

  # read in meta data
  if (min(start_n) == 1) { # Some excel files have no meta data before data table
    meta_dat <- NA
    station <- NA
  } else { # If there are meta data before data table, read it as a vector
    meta_dat <- read_excel(
      file,
      col_names = F,
      n_max = start_n[1] - 1
    ) %>% pull()
    station <- meta_dat[2] %>%
      strsplit(split = ": ") %>%
      unlist() %>%
      tail(1) # extract station name from meta data vector
  }

  location <- file %>%
    strsplit(split = c("/")) %>%
    unlist() %>%
    tail(1) %>%
    strsplit(split = " \\(") %>%
    unlist() %>%
    head(1) # city name extracted from file name

  # get coordinates of station
  if (!is.na(station)) { # use both city and station name
    station_info <- data.frame(station = station, location = location) %>%
      stringdist_inner_join(station_df, by = c("location", "station"), max_dist = 20, distance_col = "distance") %>% # join with all entries in station info, measuring dissimilarity in station and location, because their names might be slightly different in both tables
      arrange(station.distance, location.distance) %>%
      head(1) %>% # there might be several close matches. take the closest.
      rename(station = station.y) %>%
      rename(location = location.y) %>%
      dplyr::select(-station.x, -location.x, -station.distance, -location.distance, -distance)
  } else { # use only city name if station name is not available
    station_info <- data.frame(location = location) %>%
      stringdist_inner_join(station_df, by = "location", max_dist = 20, distance_col = "distance") %>%
      arrange(distance) %>%
      head(1) %>%
      rename(location = location.y) %>%
      dplyr::select(-location.x, -distance)
  }

  # Extract pollen and spore data table
  if (length(start_n) == 1) { # only pollen, no spores
    colnum <- ncol(read_excel(
      file,
      skip = start_n[1] - 1
    )) # find number of columns

    pollen_dat <- read_excel(
      file,
      skip = start_n[1] - 1,
      col_types = c(
        "date",
        rep("numeric", colnum - 1)
      )
    )
    if (is.na(pollen_dat[2, 1])) { # meaning finer taxonomic resolutions available
      genus_names <- read_excel(
        file,
        skip = start_n[1] - 1
      ) %>%
        slice(1) %>% # read the first row with finer taxonomic resolution
        gather(key = "old", value = "new") %>% # coarse and fine taxonomy into long format
        mutate(new = case_when(
          is.na(new) ~ old,
          TRUE ~ new
        )) # sometimes only coarse taxonomy is present, then keep that

      pollen_dat <- read_excel(
        file,
        skip = start_n[1] - 1,
        col_types = c(
          "date",
          rep("numeric", colnum - 1)
        ),
        col_names = genus_names$new # change to new colnames with finer taxonomic resolution
      )
    }
    pollen_dat <- pollen_dat %>%
      filter(!is.na(Date)) %>% # filter out rows that are not count data
      gather(key = "taxa", value = "count", -Date) %>% # to long format
      filter(!str_detect(taxa, "Station")) %>% # filter out additional info that are not count data, like "Station Name", "Station Postal Code", "Station State"
      filter(!taxa %in% c("Comment", "WeatherNotes")) %>%
      mutate(group = "pollen") %>%
      rowwise() %>%
      group_by(Date, taxa, group) %>%
      summarize(count = sum(count)) %>%
      ungroup()

    # combine with station info
    nab_df_list[[i]] <- pollen_dat %>%
      cbind(station_info)
  }

  if (length(start_n) == 2) { # both pollen and spores
    colnum_pollen <- ncol(read_excel(
      file,
      skip = start_n[1] - 1,
      n_max = (start_n[2] - 2) - (start_n[1] + 1)
    ))

    pollen_dat <- read_excel(
      file,
      skip = start_n[1] - 1,
      n_max = (start_n[2] - 2) - (start_n[1] + 1),
      col_types = c(
        "date",
        rep("numeric", colnum_pollen - 1)
      )
    )

    if (is.na(pollen_dat[2, 1])) { # meaning finer taxonomic resolutions available
      genus_names <- read_excel(
        file,
        skip = start_n[1] - 1,
        n_max = (start_n[2] - 2) - (start_n[1] + 1)
      ) %>%
        slice(1) %>%
        gather(key = "old", value = "new") %>%
        mutate(new = case_when(
          is.na(new) ~ old,
          TRUE ~ new
        ))

      pollen_dat <- read_excel(
        file,
        skip = start_n[1] - 1,
        n_max = (start_n[2] - 2) - (start_n[1] + 1),
        col_types = c(
          "date",
          rep("numeric", colnum_pollen - 1)
        ),
        col_names = genus_names$new
      )
    }
    pollen_dat <- pollen_dat %>%
      filter(!is.na(Date)) %>%
      gather(key = "taxa", value = "count", -Date) %>%
      filter(!str_detect(taxa, "Station")) %>%
      filter(!taxa %in% c("Comment", "WeatherNotes")) %>%
      mutate(group = "pollen") %>%
      rowwise() %>%
      group_by(Date, taxa, group) %>%
      summarize(count = sum(count)) %>%
      ungroup()

    colname_spore <- ncol(read_excel(
      file,
      skip = start_n[2] - 1
    ))
    spore_dat <- read_excel(
      file,
      skip = start_n[2] - 1,
      col_types = c(
        "date",
        rep("numeric", colname_spore - 1)
      )
    ) %>%
      filter(!is.na(Date)) %>%
      gather(key = "taxa", value = "count", -Date) %>%
      filter(!str_detect(taxa, "Station")) %>%
      filter(!taxa %in% c("Comment", "WeatherNotes")) %>%
      mutate(group = "spore")

    if (nrow(spore_dat) != 0) {
      nab_df_list[[i]] <- bind_rows(pollen_dat, spore_dat) %>%
        cbind(station_info)
    } else {
      nab_df_list[[i]] <- pollen_dat %>%
        cbind(station_info)
    }
  }
  print(i)
}

nab_df <- bind_rows(nab_df_list)

write_rds(nab_df, file = "/data/ZHULAB/phenology/nab/nab_dat.rds")
```

### Resolve taxonomy
```{r, eval=FALSE}
nab_df <- read_rds("/data/ZHULAB/phenology/nab/nab_dat.rds")

## get all distinct taxa and clean up their names
nab_taxa_df <- nab_df %>%
  distinct(taxa) %>%
  filter(!taxa %in% c("Total Pollen Count", "Total Spore Count")) %>%
  rename(taxa_raw = taxa) %>%
  rowwise() %>%
  mutate(taxa_clean = str_split(taxa_raw, pattern = "/", simplify = T)[1]) %>% # some taxa have alternative names, e.g., Chenopodiaceae/Amaranthaceae
  mutate(taxa_clean = str_split(taxa_clean, pattern = " \\(", simplify = T)[1]) %>% # e.g., Asteraceae (Excluding Ambrosia and Artemisia)
  mutate(taxa_clean = str_replace(taxa_clean, pattern = "-type", "")) %>% # e.g., Leptosphaeria-type
  mutate(taxa_clean = str_replace(taxa_clean, pattern = "Unidentified ", "")) %>% # e.g., Unidentified Fungi
  mutate(taxa_clean = str_replace(taxa_clean, pattern = "Undifferentiated ", "")) %>% # e.g., Undifferentiated Ascospores
  mutate(taxa_clean = str_replace(taxa_clean, pattern = "Other ", "")) %>% # e.g., Other Grass Pollen
  mutate(taxa_clean = str_replace(taxa_clean, pattern = "spores", "mycota")) %>% # e.g., Undifferentiated Ascospores
  mutate(taxa_clean = str_replace(taxa_clean, pattern = " ", "")) %>% # remove space
  mutate(taxa_clean = case_when(
    taxa_clean == "Rusts" ~ "Pucciniales",
    taxa_clean == "Smuts" ~ "Myxomycetes",
    taxa_clean == "Dreshslera" ~ "Helminthosporium",
    (taxa_clean == "GrassPollen" | taxa_clean == "WeedPollen") ~ "Poaceae",
    taxa_clean == "TreePollen" ~ "Tracheophyta",
    taxa_clean == "Pollen" ~ "Viridiplantae",
    TRUE ~ taxa_clean
  )) %>% # manual cleaning
  arrange(taxa_clean)

## Use taxize to resolve taxonomy
# Find sources id
# gnr_datasources()

# match with names in databases
resolve_df <- nab_taxa_df %>%
  pull(taxa_clean) %>%
  gnr_resolve(data_source_ids = c(4, 11), with_context = T, best_match_only = T, fields = "all") %>% # NCBI and GBIF databases
  full_join(nab_taxa_df,
    by = c("user_supplied_name" = "taxa_clean")
  ) %>%
  rename(taxa_clean = user_supplied_name) %>%
  mutate(same = (taxa_clean == matched_name)) # check if all taxa names are valid

# some taxa are incorrectly identified as Metazoa. Resolve again for those.
resolve_df_correct <- resolve_df %>%
  filter(str_detect(classification_path, "Metazoa")) %>%
  pull(taxa_clean) %>%
  gnr_resolve(data_source_ids = c(4, 11), best_match_only = F, fields = "all") %>% # NCBI and GBIF. Keep all matches here, not only the best match.
  filter(!str_detect(classification_path, "Animalia")) %>%
  filter(!str_detect(classification_path, "Metazoa")) %>%
  rename(taxa_clean = user_supplied_name) %>%
  group_by(taxa_clean) %>%
  slice(1) %>%
  ungroup()

# Correct previously resolved taxonomy
resolve_df <- resolve_df %>%
  filter(!str_detect(classification_path, "Metazoa")) %>%
  bind_rows(resolve_df_correct) %>%
  arrange(taxa_clean)

# make sure all taxa are resolved
# resolve_df %>% filter(!same) %>% View()

# get full classification
taxa_id_df <- resolve_df %>%
  dplyr::select(taxa_clean, data_source_id, taxon_id) %>%
  distinct(taxa_clean, .keep_all = T) %>%
  mutate(data_source = case_when(
    data_source_id == 4 ~ "ncbi",
    data_source_id == 11 ~ "gbif"
  ))

taxa_class_df <- vector(mode = "list", length = nrow(taxa_id_df))
for (i in 1:nrow(taxa_id_df)) {
  taxa_class_df[[i]] <-
    classification(taxa_id_df$taxon_id[i], db = taxa_id_df$data_source[i])  [[1]] %>%
    as_tibble() %>%
    filter(rank %in% c("kingdom", "phylum", "class", "order", "family", "genus", "species")) %>%
    mutate(rank = factor(rank, levels = c("kingdom", "phylum", "class", "order", "family", "genus", "species"))) %>%
    dplyr::select(-id) %>%
    spread(key = "rank", value = "name") %>%
    mutate(taxa_clean = taxa_id_df$taxa_clean[i])
}
taxa_class_df <- bind_rows(taxa_class_df)

nab_taxa_df <- nab_taxa_df %>%
  left_join(taxa_class_df, by = "taxa_clean") %>%
  mutate(kingdom = str_replace(kingdom, "Plantae", "Viridiplantae")) %>% # manual correction
  mutate(kingdom = case_when(
    phylum == "Oomycota" ~ "Chromista",
    TRUE ~ kingdom
  ))

write_rds(nab_taxa_df, "/data/ZHULAB/phenology/nab/nab_taxa.rds")
```

Combine nab data and taxa information. Some preprocessing.
```{r, eval=TRUE}
nab_df <- read_rds("./data/NAB/nab_dat.rds")
nab_taxa_df <- read_rds("./data/NAB/nab_taxa.rds")

nab_with_taxa_df <- nab_df %>%
  rename(taxa_raw = taxa) %>%
  left_join(nab_taxa_df, by = "taxa_raw") %>%
  rename(taxa = taxa_clean) %>%
  mutate(family = case_when(
    taxa_raw == "Total Pollen Count" ~ "Total",
    TRUE ~ family
  )) %>%
  mutate(genus = case_when(
    taxa_raw == "Total Pollen Count" ~ "Total",
    TRUE ~ genus
  )) %>%
  filter(kingdom == "Viridiplantae" | is.na(kingdom)) %>%
  group_by(Date, lat, lon, station, location, id, family, genus, taxa) %>%
  summarise(count = sum(count)) %>%
  ungroup() %>%
  mutate(date = as.Date(Date)) %>%
  dplyr::select(-Date)
```

Focus on ten cities.
* Exceptions: Denver pollen data are from Colorado Springs; Austin pollen data are from Georgetown; Detroit pollen data are from Sylvania.
```{r, eval=TRUE}
# summarize station info
meta_df <- nab_with_taxa_df %>%
  filter(taxa %in% taxa_short_list) %>% # limit to taxa studied
  drop_na(count) %>%
  group_by(station, location, lat, lon, id) %>%
  summarise(
    mindate = min(date),
    maxdate = max(date),
    n = n()
  ) %>%
  mutate(range = maxdate - mindate) %>%
  ungroup() %>%
  arrange(desc(n)) %>%
  mutate(site = NA)

# match NAB stations with cities studied
meta_df$site[meta_df$id == 26] <- "NY"
meta_df$site[meta_df$id == 7] <- "SJ"
meta_df$site[meta_df$id == 5] <- "AT"
meta_df$site[meta_df$id == 33] <- "ST"
meta_df$site[meta_df$id == 2] <- "HT"
meta_df$site[meta_df$id == 48] <- "TP"
meta_df$site[meta_df$id == 12] <- "DV"
meta_df$site[meta_df$id == 47] <- "DT"
meta_df$site[meta_df$id == 17] <- "KC"
meta_df$site[meta_df$id == 37] <- "SL"
meta_df <- meta_df %>%
  left_join(data.frame(site = site_list, sitename = sitename_list), by = "site")

write_rds(meta_df, "./data/NAB/meta_dat.rds")

# make map
p_pollen_map <- ggplot() +
  geom_polygon(data = map_data("state"), aes(x = long, y = lat, group = group), fill = "white") +
  geom_path(data = map_data("state"), aes(x = long, y = lat, group = group), color = "grey50", alpha = 0.5, lwd = 0.2) +
  theme_void() +
  geom_point(data = meta_df, aes(x = lon, y = lat), pch = 10, color = "black", cex = 3) +
  geom_label_repel(data = meta_df %>% filter(site %in% site_list), aes(x = lon, y = lat, label = sitename)) +
  geom_point(data = meta_df %>% filter(site %in% site_list), aes(x = lon, y = lat), pch = 10, color = "red", cex = 3) +
  # coord_equal()+
  coord_map("bonne", lat0 = 50)
p_pollen_map
```

View pollen phenology in study sites.
```{r, eval=TRUE}
site_lat <- meta_df %>%
  filter(site %in% site_list) %>%
  arrange(lat) %>%
  pull(sitename)

labelfunc_x <- function(x) {
  origin <- as.Date("2021-01-01")
  format(origin + x, format = "%b")
}

p_nab_ts <- nab_with_taxa_df %>%
  left_join(meta_df %>% dplyr::select(id, site, sitename), by = "id") %>%
  filter(site %in% site_list) %>%
  filter(taxa %in% taxa_short_list) %>%
  filter(!taxa %in% c("Poaceae", "Ambrosia")) %>%
  mutate(doy = format(date, "%j") %>% as.integer()) %>%
  mutate(year = format(date, "%Y") %>% as.integer()) %>%
  # filter(year %in% year_list) %>%
  # group_by(taxa, site) %>%
  mutate(count = (count + 1) %>% log(10)) %>%
  # mutate(count_st=(count-min(count, na.rm = T))/(max(count, na.rm = T)-min(count, na.rm = T))) %>%
  mutate(taxa_parse = case_when(
    !taxa %in% c("Cupressaceae", "Pinaceae", "Poaceae") ~ paste0("italic('", taxa, "')"),
    TRUE ~ taxa
  )) %>%
  # mutate(sitename=as.factor(sitename)) %>%
  mutate(sitename = fct_relevel(sitename, levels = site_lat)) %>%
  mutate(taxa_parse = fct_relevel(taxa_parse, levels = data.frame(taxa = unique(taxa_short_list)) %>%
    mutate(taxa_parse = case_when(
      !taxa %in% c("Cupressaceae", "Pinaceae", "Poaceae") ~ paste0("italic('", taxa, "')"),
      TRUE ~ taxa
    )) %>%
    pull(taxa_parse))) %>%
  ggplot() +
  geom_tile(aes(x = doy, y = sitename, fill = count), alpha = 1) +
  facet_wrap(. ~ taxa_parse, labeller = label_parsed) +
  ylab("") +
  xlab("") +
  theme_classic() +
  # scale_x_continuous(breaks = scales::pretty_breaks(n = 2)) +
  scale_x_continuous(labels = labelfunc_x) +
  theme(
    axis.line.y = element_blank(),
    # axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  ) +
  # theme(strip.text.y= element_text(angle = 0))+
  theme(
    legend.position = "bottom",
    legend.key.width = unit(2, "cm")
  ) +
  scale_fill_gradient(
    low = "light yellow", high = "red", na.value = "white",
    breaks = (c(0, 1, 10, 100, 1000, 10000) + 1) %>% log(10),
    labels = c(0, 1, 10, 100, 1000, 10000),
    name = expression(Pollen ~ concentration ~ (grains / m^3))
  )
p_nab_ts
```

## Plant location data
### Tree inventory
Read in and clean data. Each tree inventory has a different format, so they have to be processed differently.
```{r, eval=FALSE}
cl <- makeCluster(length(site_list), outfile = "")
registerDoSNOW(cl)

trees_df_list <- foreach(
  site = site_list,
  .packages = c("tidyverse", "rgdal")
) %dopar% {
  if (site == "KC" | site == "SL") {
    # Data from urban FIA
    # Download https://experience.arcgis.com/experience/3641cea45d614ab88791aef54f3a1849/page/Urban-Datamart/
    # Manual: https://www.fia.fs.fed.us/library/database-documentation/urban/dbDescription/Urban_FIADB_User_Guides_Database_Description_ver3-0_2021_03_17.pdf

    # tree table
    id_tree_df <- read_csv("./data/occurrence/StreetTrees/UrbanFIA/ID_TREE.csv") %>%
      filter(statecd == 29) %>% # Missouri
      filter(countycd %in% case_when(site == "KC" ~ 95, site == "SL" ~ c(189, 510))) %>%
      dplyr::select(plotid, id = cn, spcd, statuscd) %>%
      group_by(id) %>%
      filter(sum(statuscd != 1) == 0) %>% # living trees
      ungroup() %>%
      distinct(id, .keep_all = T) %>% # in case one tree is surveyed repeatedly
      dplyr::select(-statuscd)

    # plot table
    id_plot_df <- read_csv("./data/occurrence/StreetTrees/UrbanFIA/ID_PLOT.csv") %>%
      dplyr::select(plotid, lat, lon)

    # species reference table
    ref_sp_df <- read_csv("./data/occurrence/StreetTrees/UrbanFIA/REF_SPECIES.csv") %>%
      dplyr::select(spcd, genus, species) %>%
      mutate(species = paste(genus, species))

    # join tables
    trees_df <- id_tree_df %>%
      left_join(id_plot_df, by = "plotid") %>%
      left_join(ref_sp_df, by = "spcd") %>%
      dplyr::select(-plotid, -spcd) %>%
      mutate(site = site)
  }
  if (site == "DT") {
    # Data from Dan Katz
    trees_df <- read_csv("./data/occurrence/StreetTrees/Tree_Inventory_Detroit.csv") %>%
      distinct(id = ID, species = SPP, x=POINT_X, y=POINT_Y)
    # reproject to WGS84
    pts <- SpatialPoints(trees_df[, c("x", "y")],
      proj4string = CRS("ESRI:102290")
    )
    # ESRI:102290: NAD 1983 HARN StatePlane Michigan South FIPS 2113
    # https://spatialreference.org/ref/?search=michigan
    
    # trees_df <- read_csv("./data/Detroit_oak_pheno_obs_spring_2017.csv") %>%
    #   distinct(id = tree, species = Species, x, y)

    # reproject to WGS84
    # pts <- SpatialPoints(trees_df[, c("x", "y")],
    #   proj4string = CRS("+init=EPSG:3857")
    # )
    # +proj=merc +a=6378137 +b=6378137 +lat_ts=0.0 +lon_0=0.0 +x_0=0.0 +y_0=0 +k=1.0 +units=m +nadgrids=@null +wktext +no_defs
    
    pts_reproj <- spTransform(
      pts,
      CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
    )

    trees_df <- cbind(trees_df %>% dplyr::select(-x, -y), coordinates(pts_reproj)) %>%
      rename(lat = y, lon = x) %>%
      as_tibble() %>%
      mutate(site = site)
  }
  if (site == "DV") {
    # Data from OpenTrees.org
    trees_df <- read_csv("./data/occurrence/StreetTrees/Tree_Inventory_Denver.csv") %>%
      dplyr::select(id = SITE_ID, species = SPECIES_BO, time = INVENTORY_DATE, lat = Y_LAT, lon = X_LONG) %>%
      arrange(desc(time)) %>%
      distinct(id, species, .keep_all = T) %>%
      dplyr::select(-time) %>%
      filter(lon != 0) %>%
      mutate(site = site)
  }
  if (site == "TP") {
    # Data from https://www.opentreemap.org/tampa/map/
    trees_df <- read_csv("./data/occurrence/StreetTrees/Tree_Inventory_Tampa.csv") %>%
      dplyr::select(id = `Tree Id`, genus = Genus, species = Species, lat = `Point Y`, lon = `Point X`) %>% # checked that there is no repeated tree id
      mutate(species = paste0(genus, " ", species)) %>%
      dplyr::select(-genus) %>%
      mutate(site = site)
  }
  if (site == "HT") {
    # Data from  https://koordinates.com/layer/25245-houston-texas-street-tree-inventory/data/
    trees_df <- read_csv("./data/occurrence/StreetTrees/Tree_Inventory_Houston/houston-texas-street-tree-inventory.csv") %>%
      dplyr::select(species_common = Species, X = Shape_X, Y = Shape_Y) %>%
      rowwise() %>%
      mutate(common = case_when(
        str_detect(species_common, ", spp.") ~ str_replace(species_common, ", spp.", ""), # for coarse common names, e.g., "Oak, spp."
        (str_detect(species_common, ", ") & !str_detect(species_common, ", spp.")) ~ paste0(str_split(species_common, ", ")[[1]][2], " ", str_split(species_common, ", ")[[1]][1]), # reverse order of common name, e.g., "Oak, Water"
        TRUE ~ species_common
      )) %>%
      ungroup() %>%
      distinct(X, Y, .keep_all = T) %>% # in case same tree is sampled repeatedly
      mutate(id = row_number())

    # reproject to WGS84
    pts <- SpatialPoints(trees_df[, c("X", "Y")],
      proj4string = CRS("+proj=merc +a=6378137 +b=6378137 +lat_ts=0 +lon_0=0 +x_0=0 +y_0=0 +k=1 +units=m +nadgrids=@null +wktext +no_defs")
    )
    pts_reproj <- spTransform(
      pts,
      CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
    )

    trees_df <- cbind(trees_df %>% dplyr::select(-X, -Y), coordinates(pts_reproj)) %>%
      rename(lat = Y, lon = X)

    # Use taxize to match common name with scientific name
    if (!file.exists("./data/occurrence/StreetTrees/Tree_Inventory_Houston/species_reference.csv")) {
      id2comm <- function(pageid) {
        common_matches <- eol_pages(taxonconceptID = pageid, common_names = TRUE)$vernacular
        if (is.null(common_matches)) {
          common_match <- ""
        } else {
          common_matches <- common_matches %>% filter(language == "en")
          if (nrow(common_matches) > 0) {
            common_match <- common_matches %>%
              pull(vernacularname) %>%
              table() %>%
              sort(decreasing = TRUE) %>%
              head(1) %>%
              names()
          }
        }
        return(common_match)
      }

      comm2sci_new <- function(common) {
        species <- comm2sci(common, simplify = T)[[1]]
        if (length(species) == 0) {
          res <- comm2sci(common, db = "eol", simplify = F)[[1]]
          if (length(res) == 0) {
            species <- NA
          } else {
            species <- res %>%
              as_tibble() %>%
              rowwise() %>%
              mutate(common_match = id2comm(pageid)) %>%
              ungroup() %>%
              stringdist_inner_join(data.frame(common_name = common), by = c("common_match" = "common_name"), max_dist = 20, distance_col = "distance") %>%
              arrange(distance, pageid) %>%
              head(1) %>%
              filter(common_match != "") %>%
              pull(name)
            if (length(species) == 0) {
              species <- NA
            }
          }
          print(paste(common, species))
        }

        if (is.na(species)) {
          species <- readline(prompt = paste0(species_df$common[i], ". Manually enter scientific name: "))
        }

        return(species)
      }

      species_df <- trees_df %>%
        distinct(common) %>%
        rowwise() %>%
        mutate(species = comm2sci_new(common))
      write_csv(species_df, "./data/occurrence/StreetTrees/Tree_Inventory_Houston/species_reference.csv")
    } else {
      species_df <- read_csv("./data/occurrence/StreetTrees/Tree_Inventory_Houston/species_reference.csv")
    }

    trees_df <- trees_df %>%
      left_join(species_df, by = "common") %>%
      dplyr::select(-species_common, -common) %>%
      mutate(site = site)
  }
  if (site == "NY") {
    trees_df <- read_csv("./data/occurrence/StreetTrees/Tree_Inventory_NewYork.csv") %>%
      dplyr::select(id = tree_id, species = spc_latin, lat = latitude, lon = longitude) %>% # checked that there is no repeated tree id
      mutate(site = site)
  }
  if (site == "AT") {
    trees_df <- read_csv("./data/occurrence/StreetTrees/Tree_Inventory_Austin/Tree_Inventory_Austin.csv") %>%
      dplyr::select(id = OBJECTID, species = SPECIES, coordinates = the_geom) %>% # checked that there is no repeated tree id
      mutate(coordinates = str_replace(coordinates, "POINT \\(", "")) %>%
      mutate(coordinates = str_replace(coordinates, "\\)", "")) %>%
      rowwise() %>%
      mutate(
        lon = str_split(coordinates, pattern = " ", simplify = T)[1],
        lat = str_split(coordinates, pattern = " ", simplify = T)[2]
      ) %>%
      ungroup() %>%
      mutate(
        lon = as.numeric(lon),
        lat = as.numeric(lat)
      ) %>%
      dplyr::select(-coordinates) %>%
      mutate(site = site)
  }
  if (site == "SJ") {
    trees_df <- read_csv("./data/occurrence/StreetTrees/Tree_Inventory_SanJose/Tree_Inventory_SanJose.csv") %>%
      dplyr::select(id = OBJECTID, species = NAMESCIENTIFIC, Y = Y, X = X) # checked that there is no repeated tree id

    # shape <- readOGR(dsn = "/data/ZHULAB/phenology/occurrence/StreetTrees/Tree_Inventory_SanJose/Street_Tree.shp")
    # proj4string(shape)
    projection_sj <- "+proj=lcc +lat_0=36.5 +lon_0=-120.5 +lat_1=38.4333333333333 +lat_2=37.0666666666667 +x_0=2000000.0001016 +y_0=500000.0001016 +datum=NAD83 +units=us-ft +no_defs"
    pts <- SpatialPoints(trees_df[, c("X", "Y")],
      proj4string = CRS(projection_sj)
    )
    pts_reproj <- spTransform(
      pts,
      CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
    )

    trees_df <- cbind(trees_df %>% dplyr::select(-X, -Y), coordinates(pts_reproj)) %>%
      rename(lat = Y, lon = X) %>%
      mutate(site = site)
  }
  if (site == "ST") {
    trees_df <- read_csv("./data/occurrence/StreetTrees/Tree_Inventory_Seattle.csv") %>%
      dplyr::select(id = OBJECTID, species = SCIENTIFIC_NAME, lat = Y, lon = X) %>% # checked that there is no repeated tree id
      mutate(site = site)
  }
  trees_df
}

stopCluster(cl)

trees_df <- bind_rows(trees_df_list) %>%
  rowwise() %>%
  mutate(genus = str_split(species, pattern = " ", simplify = T)[1]) %>% # get genus name from species name
  ungroup() %>%
  distinct(id, site, .keep_all = T) %>% # in case one tree is surveyed repeatedly
  mutate(genus_id = as.integer(as.factor(genus)))
```

Find family names from genus names. This step needs supervision.
```{r, eval=FALSE}
keywords<-c("unknown", "vacant", "tbd", "na", "x", "other")
key_regex<-regex(paste("\\b(?i)", keywords, "\\b", sep = "", collapse = "|"))
genus_to_family <- trees_df %>%
  distinct(genus) %>%
  mutate(genus=str_replace_all(genus, "[^[:alnum:]]", "")) %>% 
  mutate(genus=str_replace_all(genus, key_regex, "")) %>% 
  drop_na() %>% 
  filter(genus!="") %>% 
  mutate(family=NA)
for (i in 1:nrow(genus_to_family)) {
  family<- tax_name(sci = genus_to_family$genus[i], get = "family", messages = F, db = "ncbi")$family
  genus_to_family$family[i]<-family
  Sys.sleep(0.5)
}
  
write_rds(genus_to_family, "./data/occurrence/genus_to_family.rds")
```

Join with inventory data and save.
```{r, eval=FALSE}
genus_to_family <- read_rds("./data/occurrence/genus_to_family.rds")

trees_df <- trees_df %>%
  left_join(genus_to_family, by = "genus")
write_rds(trees_df, "./data/occurrence/street_trees.rds")
```

### Grass patch
```{r, eval=FALSE}
# Land cover data manually downloaded from Sentinel-2 Land Use/ Land Cover Downloader
# https://www.arcgis.com/apps/instant/media/index.html?appid=fc92d38533d440078f17678ebc20e8e2
grass_path <- "/data/ZHULAB/phenology/occurrence/LULC/"
trees_df <- read_rds("/data/ZHULAB/phenology/occurrence/street_trees.rds")

cl <- makeCluster(length(site_list), outfile = "")
registerDoSNOW(cl)

grass_df_city_list <- foreach(
  siteoi = site_list,
  .packages = c("tidyverse", "raster")
) %dopar% {

  # get extent from tree inventory
  trees_df_city <- trees_df %>%
    filter(site == siteoi) %>%
    drop_na(lon, lat)
  bbox <- extent(min(trees_df_city$lon), max(trees_df_city$lon), min(trees_df_city$lat), max(trees_df_city$lat))

  grass_df_city_year_list <- vector(mode = "list")
  for (yearoi in year_list) {
    # get file name(s) for each site and year
    files <- list.files(paste0(grass_path, siteoi), pattern = paste0(yearoi, "0101-"), full.names = T)

    # read raster(s)
    ras <- raster(files)

    # crop to city
    bbox_sp <- as(bbox, "SpatialPolygons")
    projection(bbox_sp) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
    bbox_reproj <- spTransform(bbox_sp, proj4string(ras))
    ras_cr <- crop(ras, bbox_reproj)

    # get coordinates of grass
    grass_df_year <- as.data.frame(ras_cr, xy = T) %>%
      `colnames<-`(c("x", "y", "class")) %>%
      filter(class == 11) %>%
      dplyr::select(-class) %>%
      mutate(year = yearoi)
    grass_df_city_year_list[[yearoi %>% as.character()]] <- grass_df_year
    print(paste0(siteoi, yearoi))
  }
  grass_df_city <- bind_rows(grass_df_city_year_list) %>%
    mutate(grass = 1) %>% # choose pixels that are grass in all years
    spread(key = "year", value = "grass") %>%
    drop_na() %>%
    dplyr::select(x, y) %>%
    sample_n(min(10000, nrow(.))) # subset when there are too many grass pixels

  # reproject
  grass_sp_reproj <- SpatialPoints(grass_df_city[, c("x", "y")],
    proj4string = CRS(proj4string(ras))
  )
  grass_sp <- spTransform(
    grass_sp_reproj,
    CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
  )

  # get coordinates
  grass_df_city <- coordinates(grass_sp) %>%
    as_tibble() %>%
    `colnames<-`(c("lon", "lat")) %>%
    mutate(site = siteoi) %>%
    mutate(id = row_number()) %>%
    mutate(family = "Poaceae") %>%
    mutate(genus = "Unknown") %>%
    mutate(genus_id = 999)
  grass_df_city
}
stopCluster(cl)

grass_df <- bind_rows(grass_df_city_list)
write_rds(grass_df, "/data/ZHULAB/phenology/occurrence/grass.rds")
```

### Ragweed observation
Get ragweed occurrence information from GBIF.
```{r, eval=FALSE}
trees_df <- read_rds("/data/ZHULAB/phenology/occurrence/street_trees.rds")

cl <- makeCluster(length(site_list), outfile = "")
registerDoSNOW(cl)

ragweed_df_city_list <- foreach(
  siteoi = site_list,
  .packages = c("tidyverse", "raster", "sf", "spocc")
) %dopar% {
  # get extent
  trees_df_city <- trees_df %>%
    filter(site == siteoi) %>%
    drop_na(lon, lat)
  bbox <- extent(min(trees_df_city$lon), max(trees_df_city$lon), min(trees_df_city$lat), max(trees_df_city$lat))
  # make it a polygon
  bbox_sp <- as(bbox, "SpatialPolygons")
  projection(bbox_sp) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"

  # get gbif data
  res <- occ(
    query = "Ambrosia", from = "gbif", has_coords = TRUE, limit = 1e6,
    geometry = st_bbox(bbox_sp),
    date = c(as.Date("2018-01-01"), as.Date("2021-12-31")),
    gbifopts = list(
      hasGeospatialIssue = FALSE
    )
  )

  # get coordinates
  ragweed_df_city <- res$gbif$data[[1]] %>%
    dplyr::select(lon = longitude, lat = latitude, species) %>%
    mutate(site = siteoi) %>%
    mutate(family = "Asteraceae") %>%
    mutate(genus = "Ambrosia") %>%
    mutate(genus_id = 998)
  ragweed_df_city
}
stopCluster(cl)

ragweed_df <- bind_rows(ragweed_df_city_list)
write_rds(ragweed_df, "/data/ZHULAB/phenology/occurrence/ragweed.rds")
```

### View plant occurrence data

Put tree, grass, and ragweed data in one data frame.
```{r, eval=TRUE}
trees_df <- read_rds("./data/occurrence/street_trees.rds")
# test<- trees_df %>% filter(site=="AT") %>% sample_n(1)
# paste0(test$lat,",",test$lon, ",", test$species)
grass_df <- read_rds("./data/occurrence/grass.rds")
ragweed_df <- read_rds("./data/occurrence/ragweed.rds")
plant_df <- bind_rows(trees_df, grass_df, ragweed_df)
```

Map relative position of tree inventory and nab station.
```{r, eval=TRUE}
p_nab_plant_map <- ggplot() +
  geom_polygon(data = map_data("state"), aes(x = long, y = lat, group = group), fill = "white") +
  geom_path(data = map_data("state"), aes(x = long, y = lat, group = group), color = "grey50", alpha = 0.5, lwd = 0.2) +
  theme_void() +
  geom_point(
    data = plant_df %>%
      filter(site %in% site_list) %>% 
      filter(genus %in% taxa_short_list | family %in% taxa_short_list) %>%
      mutate(taxa = case_when(
        family %in% c("Poaceae", "Cupressaceae", "Pinaceae") ~ family,
        TRUE ~ genus
      )) %>%
      filter(!taxa %in% c("Poaceae", "Ambrosia")) %>%
      group_by(site) %>%
      summarize(
        midlon = median(lon, na.rm = T),
        midlat = median(lat, na.rm = T)
      ) %>%
      # sample_n(100) %>%
      ungroup(),
    aes(x = midlon, y = midlat), col = "dark blue", cex = 3, alpha = 1, pch = 0
  ) +
  geom_point(data = meta_df, aes(x = lon, y = lat), cex = 3, pch = 10, col = "black", alpha = 0.5) +
  geom_point(data = meta_df %>% filter(site %in% site_list), aes(x = lon, y = lat), cex = 3, pch = 10, col = "red", alpha = 1) +
  geom_label_repel(
    data = plant_df %>%
      filter(site %in% site_list) %>% 
      filter(genus %in% taxa_short_list | family %in% taxa_short_list) %>%
      mutate(taxa = case_when(
        family %in% c("Poaceae", "Cupressaceae", "Pinaceae") ~ family,
        TRUE ~ genus
      )) %>%
      # filter(!taxa%in% c("Poaceae", "Ambrosia")) %>%
      group_by(site) %>%
      summarize(
        midlon = median(lon, na.rm = T),
        midlat = median(lat, na.rm = T)
      ) %>%
      # sample_n(100) %>%
      ungroup() %>%
      filter(site %in% site_list) %>%
      left_join(meta_df %>% dplyr::select(site, sitename) %>% drop_na(),
        by = "site"
      ),
    aes(x = midlon, y = midlat, label = sitename), nudge_x = 1.5, nudge_y = 1.5, col = "dark blue"
  ) +
  coord_map("bonne", lat0 = 50)
p_nab_plant_map
```

Calculate distance from plants to NAB stations in the unit of km.
```{r, eval=FALSE}
distance_df <- plant_df %>%
  filter(genus %in% taxa_short_list | family %in% taxa_short_list) %>%
  left_join(meta_df %>% dplyr::select(site, sitelon = lon, sitelat = lat, sitename) %>% drop_na(), by = "site") %>%
  rowwise() %>%
  mutate(distance = distm(c(lon, lat), c(sitelon, sitelat), fun = distHaversine) %>% as.numeric() %>% `/`(1000)) %>% # distance in the unit of km
  ungroup() %>%
  group_by(site, sitename) %>%
  summarise(
    mindist = min(distance),
    maxdist = max(distance),
    meandist = mean(distance)
  )
write_rds(distance_df, "/data/ZHULAB/phenology/occurrence/distance_from_plants_to_nab_stations.rds")
```

```{r, eval=TRUE}
read_rds("./data/occurrence/distance_from_plants_to_nab_stations.rds")
```
Prepare street map as basemap. Street shapefiles for major cities manually downloaded from https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/CUWWYJ.
Boeing, Geoff, 2017, "U.S. Street Network Shapefiles, Node/Edge Lists, and GraphML Files", https://doi.org/10.7910/DVN/CUWWYJ, Harvard Dataverse, V2
```{r, eval=FALSE}

roads_files_all <- list.files("/data/ZHULAB/phenology/occurrence/roads/", pattern = "edges.shp", recursive = T, full.names = T)

cl <- makeCluster(length(site_list), outfile = "")
registerDoSNOW(cl)
roads_fort_list <-
  foreach(
    s = 1:length(site_list),
    .packages = c("rgdal", "rgeos", "stringr", "tidyverse", "raster")
  ) %dopar% {

    # read shapefile for the site
    siteoi <- site_list[s]
    sitename <- sitename_list[s]
    roads_file <- str_subset(roads_files_all, pattern = sitename %>% str_replace(" ", "_"))
    roads <- readOGR(dsn = roads_file)

    # get boundary from plant occurrence data frame
    plant_df_site <- plant_df %>% filter(site == siteoi)
    area_site <- extent(min(plant_df_site$lon) - 0.005, max(plant_df_site$lon) + 0.005, min(plant_df_site$lat) - 0.005, max(plant_df_site$lat) + 0.005)
    area_site <- as(area_site, "SpatialPolygons")
    proj4string(area_site) <- CRS(proj4string(roads))

    # clip the shapefile
    rgeos::set_RGEOS_CheckValidity(2L)
    roads_crop <- gIntersection(roads, area_site, byid = TRUE)

    # transform into a format that can be used in ggplot
    roads_sldf <- SpatialLinesDataFrame(roads_crop, data = data.frame(value = rep(1, length(roads_crop))), match.ID = F)
    roads_fort <- roads_sldf %>%
      fortify() %>%
      mutate(
        site = siteoi,
        sitename = sitename
      )

    roads_fort
  }
roads <- bind_rows(roads_fort_list)
stopCluster(cl)
write_rds(roads, "/data/ZHULAB/phenology/occurrence/roads/roads_cities.rds")
```

Map plant occurrence in each city.
```{r, eval=TRUE, fig.width=12, fig.height=30}
roads <- read_rds("./data/occurrence/roads/roads_cities.rds")

extent_df <- plant_df %>%
  group_by(site) %>%
  summarise(
    minlon = min(lon),
    minlat = min(lat),
    maxlon = max(lon),
    maxlat = max(lat)
  )

p_plant_map <- ggplot() +
  theme_void() +
  geom_line(
    data = roads %>%
      filter(site %in% c("NY")) %>%
      left_join(extent_df, by = "site") %>%
      group_by(site) %>%
      mutate(
        long = (long - (maxlon + minlon) / 2) / max((maxlon - minlon), (maxlat - minlat)),
        lat = (lat - (maxlat + minlat) / 2) / max((maxlon - minlon), (maxlat - minlat))
      ) %>%
      ungroup(),
    aes(x = long, y = lat, group = group), size = .2, alpha = 0.5
  ) +
  geom_point(
    data = plant_df %>%
      filter(site %in% c("NY")) %>%
      filter(genus %in% taxa_short_list | family %in% taxa_short_list) %>%
      mutate(taxa = case_when(
        family %in% c("Poaceae", "Cupressaceae", "Pinaceae") ~ family,
        TRUE ~ genus
      )) %>%
      filter(!taxa %in% c("Poaceae", "Ambrosia")) %>%
      group_by(taxa, site) %>%
      sample_n(min(100, n())) %>%
      ungroup() %>%
      left_join(meta_df %>% dplyr::select(site, sitename), by = "site") %>%
      left_join(extent_df, by = "site") %>%
      group_by(site) %>%
      mutate(
        lon = (lon - (maxlon + minlon) / 2) / max((maxlon - minlon), (maxlat - minlat)),
        lat = (lat - (maxlat + minlat) / 2) / max((maxlon - minlon), (maxlat - minlat))
      ) %>%
      ungroup(),
    aes(x = lon, y = lat, col = taxa), alpha = 0.3
  ) +
  facet_wrap(. ~ sitename, ncol = 1) +
  guides(col = guide_legend(title = "Taxa")) +
  theme(legend.position = "bottom") +
  # ylab("Latitude")+
  # xlab("Longitude")+
  coord_equal()
p_plant_map
```

## NPN data
```{r, eval=FALSE}
phenophases <- npn_phenophases()
species_list <- npn_species()

# download all NPN data for taxa studied
npn_path <- "/data/ZHULAB/phenology/NPN/"

cl <- makeCluster(length(site_list), outfile = "")
registerDoSNOW(cl)

npn_df_list <- vector(mode = "list")
for (taxaoi_short in taxa_short_list %>% unique()) {
  if (!file.exists(paste0(npn_path, taxaoi_short, ".rds"))) {
    spid <- species_list %>%
      filter(genus == taxaoi_short | family_name == taxaoi_short) %>%
      pull(species_id)

    npn_data <- npn_download_status_data(request_source = "YS", years = c(2000:2022), species_id = spid)

    write_rds(npn_data, paste0(npn_path, taxaoi_short, ".rds"))
  } else {
    npn_data <- read_rds(paste0(npn_path, taxaoi_short, ".rds"))
  }

  npn_taxa_df_list <- foreach(
    siteoi = site_list,
    .packages = c("tidyverse", "geosphere")
  ) %dopar% {
    lat_oi <- meta_df %>%
      filter(site == siteoi) %>%
      pull(lat) %>%
      mean()
    lon_oi <- meta_df %>%
      filter(site == siteoi) %>%
      pull(lon) %>%
      mean()

    npn_flower_df <- npn_data %>%
      filter(phenophase_status == 1) %>%
      filter(phenophase_description %in% c("Full pollen release (conifers)", "Pollen release (conifers)", "Pollen cones (conifers)", "Open pollen cones (conifers)", "Full flowering (50%)", "Flowers or flower buds", "Pollen release (flowers)"))

    if (nrow(npn_flower_df > 0)) {
      npn_flower_df <- npn_flower_df %>%
        rowwise() %>%
        mutate(distance = distm(c(longitude, latitude), c(lon_oi, lat_oi), fun = distHaversine) %>% as.numeric()) %>% # distance in the unit of m
        ungroup() %>%
        filter(distance <= 500000) %>% # within 500 km of the NAB station
        dplyr::select(date = observation_date, lon = longitude, lat = latitude, doy = day_of_year) %>%
        mutate(date = as.Date(date)) %>%
        mutate(site = siteoi)
    } else {
      npn_flower_df <- data.frame(lon = numeric(0), lat = numeric(0), doy = integer(0), date = character(0), site = character(0)) %>% mutate(date = as.Date(date))
    }

    npn_count_df <- npn_flower_df %>%
      group_by(site, date) %>%
      summarise(count = n()) %>%
      ungroup()

    print(paste0(taxaoi_short, ", ", siteoi))
    npn_count_df
  }

  npn_df_list[[taxaoi_short]] <- bind_rows(npn_taxa_df_list) %>%
    mutate(taxa = taxaoi_short)
}
stopCluster(cl)
npn_df <- bind_rows(npn_df_list)
write_rds(npn_df, paste0(npn_path, "npn_dat.rds"))
```

View flowering phenology (from NPN data) in study sites.
```{r, eval=TRUE}
npn_path <- "./data/NPN/"
npn_df_all <- read_rds(paste0(npn_path, "npn_dat.rds"))
npn_df_all %>%
  left_join(meta_df %>% dplyr::select(id, site, sitename), by = "site") %>%
  filter(site %in% site_list) %>%
  filter(taxa %in% taxa_short_list) %>%
  mutate(doy = format(date, "%j") %>% as.integer()) %>%
  mutate(year = format(date, "%Y") %>% as.integer()) %>%
  # filter(year %in% year_list) %>%
  group_by(taxa, site) %>%
  mutate(count_st = (count - min(count, na.rm = T)) / (max(count, na.rm = T) - min(count, na.rm = T))) %>%
  ggplot() +
  geom_point(aes(x = doy, y = count_st, group = taxa, col = taxa), alpha = 0.3) +
  facet_grid(cols = vars(taxa), rows = vars(sitename), scales = "free_y") +
  theme_classic()
```

## CHELSA data
```{r, eval=FALSE}
# Downloaded from: https://chelsa-climate.org/bioclim/
# Documentation: https://chelsa-climate.org/wp-admin/download-page/CHELSA_tech_specification_V2.pdf

chelsa_path <- "/data/ZHULAB/phenology/CHELSA/"
# read in raster
tmean_ras <- raster(paste0(chelsa_path, "bio1.tif"))
ppt_ras <- raster(paste0(chelsa_path, "bio12.tif"))
vpdmax_ras <- raster(paste0(chelsa_path, "vpd_max.tif"))

# sites as points
meta_sf <- meta_df %>%
  drop_na(site) %>%
  dplyr::select(site, lon, lat) %>%
  st_as_sf(
    coords = c("lon", "lat"),
    crs = 4326
  )

# extract chelsa data at points
chelsa_df <- meta_sf %>%
  mutate(
    mat = raster::extract(tmean_ras, meta_sf),
    tap = raster::extract(ppt_ras, meta_sf),
    vpd = raster::extract(vpdmax_ras, meta_sf)
  ) %>%
  as_tibble() %>%
  dplyr::select(-geometry)

write_rds(chelsa_df, "./data/chelsa.rds")
```

```{r, eval=TRUE}
chelsa_df <- read_rds("./data/chelsa.rds")
```

## PlanetScope data
### Order

```{r, eval=FALSE, echo=FALSE}
# Modify a function to record order name
planet_order_download_new <- function(order_id, order_name, api_key, order_num, overwrite_opt = FALSE) {
  # GET order for download
  # If you lose the order_id, don't redo the request, log onto planet and find it in the orders menu
  # order_id for example SMV2 order: "dab92990-ce3a-456c-8ad6-ca0c569b4a1a"
  url2 <- paste0("https://api.planet.com/compute/ops/orders/v2/", order_id)

  get_order <- httr::GET(
    url = url2,
    username = api_key
  )
  # Download links are in here, under _links>results>location
  get_content <- httr::content(get_order)
  # When state = 'success', ready for download

  # check if order is ready
  while (get_content$state != "success") {
    print("Order still being proccessed, trying again in 60 seconds...")
    print(get_content$state)
    Sys.sleep(60)
    get_order <- httr::GET(url = url2, username = api_key)
    get_content <- httr::content(get_order)
  }

  ## Time to download!
  print("Starting download")

  # First create download folder:
  dir.create(order_name, showWarnings = F)

  # Download each item in order
  for (i in 1:length(get_content$`_links`$results)) {
    print(paste0("Order ", order_num, ", Download: ", round(100 * (
      i / length(get_content$`_links`$results)
    ), 2), "%"))
    # find item names in order contents
    name <- get_content$`_links`$results[[i]]$name
    findslash <- gregexpr("/", name)
    startchar <- findslash[[1]][length(findslash[[1]])] + 1
    filename <- substr(name, startchar, nchar(name))

    download_url <- get_content$`_links`$results[[i]]$location

    try({
      httr::RETRY(
        "GET",
        url = download_url,
        username = api_key,
        write_disk(
          path = paste(order_name, filename, sep = "/"),
          overwrite = overwrite_opt
        )
      )
    })
  }

  print(paste0("Download complete"))
  print(paste0("Items located in ", order_name))
}
```

User settings.
```{r, eval=TRUE}
# Set API
api_key = "REMOVED" #ysong67
# api_key <- "REMOVED" # xcui12

# Metadata filters
cloud_lim <- 0.99 # percent from 0-1
item_name <- "PSScene4Band"
# PSOrthoTile, PSScene3Band, PSScene4Band, Sentinel2L1C
# (see https://developers.planet.com/docs/data/items-assets/)
product <- "analytic_sr"
# analytic_b1, analytic_b2
# (see https://developers.planet.com/docs/data/items-assets/)
# https://developers.planet.com/docs/apis/data/psscene3-4band-deprecation/

ps_path <- "./data/PS/"
```

Order data. Orders may fail. May have to order again in that case.
```{r, eval=FALSE}
for (siteoi in site_list) {
  ps_path_site <- paste0(ps_path, siteoi, "/")
  dir.create(ps_path_site, recursive = T)
  dir.create(paste0(ps_path_site, "orders/"), recursive = T)

  # Set AOI (many ways to set this!) ultimately just need an extent()
  plant_df_site <- plant_df %>%
    filter(site == siteoi) %>%
    drop_na(lon, lat)
  bbox <- extent(min(plant_df_site$lon), max(plant_df_site$lon), min(plant_df_site$lat), max(plant_df_site$lat))

  for (year_download in 2017:2022) {
    order_df <- data.frame(year = integer(0), month = integer(0), id = character(0))
    if (year_download == 2022) {
      month_end <- 4
    } else {
      month_end <- 12
    }
    for (month_download in 1:month_end) {
      # Date range of interest
      start_year <- year_download
      end_year <- year_download
      date_start <- lubridate::floor_date(as.Date(paste0(year_download, "-", str_pad(month_download, 2, pad = "0"), "-01")), unit = "month")
      date_end <- lubridate::ceiling_date(as.Date(paste0(year_download, "-", str_pad(month_download, 2, pad = "0"), "-01")), unit = "month") - 1
      start_doy <- as.numeric(format(date_start, "%j"))
      end_doy <- as.numeric(format(date_end, "%j"))

      # Create order name
      order_name <- paste(siteoi, item_name, product, start_year, end_year, start_doy, end_doy, sep = "_")

      # Planet Orders API
      order_id <- planetR:::planet_order_request(
        api_key = api_key,
        bbox = bbox,
        date_start = date_start,
        date_end = date_end,
        start_doy = start_doy,
        end_doy = end_doy,
        cloud_lim = cloud_lim,
        item_name = item_name,
        product = product,
        order_name = order_name
      )

      # when there are too many items in an order, order id will not be returned. split order into a few groups in that case.
      if (!is.null(order_id)) {
        # store order name
        order_df <- order_df %>%
          bind_rows(data.frame(year = year_download, month = month_download, id = order_id))
      } else {
        item_num <- nrow(planet_search(
          bbox, start_doy, end_doy, date_end,
          date_start, cloud_lim, item_name, api_key
        ))
        group_num <- ceiling(item_num / 500)
        split_num <- ceiling(30 / group_num)
        doy_group <- split(seq(start_doy, end_doy), floor(seq(1:(end_doy - start_doy + 1)) / split_num))
        for (g in 1:length(doy_group)) {
          order_id <- planetR:::planet_order_request(
            api_key = api_key,
            bbox = bbox,
            date_start = date_start,
            date_end = date_end,
            start_doy = min(doy_group[[g]]),
            end_doy = max(doy_group[[g]]),
            cloud_lim = cloud_lim,
            item_name = item_name,
            product = product,
            order_name = order_name
          )
          if (!is.null(order_id)) {
            order_df <- order_df %>%
              bind_rows(data.frame(year = year_download, month = month_download, id = order_id))
          }
        }
      }

      print(paste(year_download, ", ", month_download))
    }
    dir.create(paste0(ps_path_site, "orders/"))
    write_rds(order_df, paste0(ps_path_site, "orders/", "order_", year_download, ".rds"))
  }
}
```

### Download
```{r, eval=FALSE}
for (siteoi in site_list) {
  ps_path_site <- paste0(ps_path, siteoi, "/")
  for (year_download in 2017:2022) {
    order_df <- read_rds(paste0(ps_path_site, "orders/", "order_", year_download, ".rds"))
    cl <- makeCluster(nrow(order_df), outfile = "")
    registerDoSNOW(cl)
    foreach(
      i = 1:nrow(order_df),
      .packages = c("stringr", "planetR", "httr")
    ) %dopar% {

      # Get order id
      month_download <- order_df$month[i]
      order_id <- order_df$id[i]

      # Date range of interest
      start_year <- year_download
      end_year <- year_download
      date_start <- lubridate::floor_date(as.Date(paste0(year_download, "-", str_pad(month_download, 2, pad = "0"), "-01")), unit = "month")
      date_end <- lubridate::ceiling_date(as.Date(paste0(year_download, "-", str_pad(month_download, 2, pad = "0"), "-01")), unit = "month") - 1
      start_doy <- as.numeric(format(date_start, "%j"))
      end_doy <- as.numeric(format(date_end, "%j"))

      # Set/Create Export Folder
      order_name <- paste(siteoi, item_name, product, start_year, end_year, start_doy, end_doy, sep = "_")
      exportfolder <- paste0(ps_path_site, order_name)
      dir.create(exportfolder, recursive = T, showWarnings = F)

      # Download
      Sys.sleep(i * 0.5) # Otherwise sending request to API at the same time may cause error
      planet_order_download_new(order_id, exportfolder, api_key = api_key, order_num = i, overwrite_opt = FALSE)

      print(paste(year_download, ", ", month_download))
    }

    stopCluster(cl)
  }
}
```

### Retrieve time series at plant locations
```{r, eval=FALSE}
cl <- makeCluster(50, outfile = "")
registerDoSNOW(cl)

iscomplete <- F
while (!iscomplete) { # restart when there is error, usually because of cluster connection issues
  iserror <- try(
    for (taxaoi in taxa_list) {
      taxaoi_short <- str_split(taxaoi, " ", simplify = T)[1]
      for (s in 1:length(site_list)) {
        # get plant locations
        siteoi <- site_list[s]
        plant_taxa_df <- plant_df %>%
          filter(site == siteoi) %>%
          filter(genus == taxaoi_short | family == taxaoi_short) %>%
          mutate(id = row_number()) %>%
          drop_na(lon, lat)

        # skip when there are too few plants
        if (taxaoi_short %in% c("Ambrosia", "Poaceae")) {
          min_sample_size <- 1
        } else {
          min_sample_size <- 1
        }

        if (nrow(plant_taxa_df) >= min_sample_size) {
          # plants as points
          plant_taxa_sp <- SpatialPoints(plant_taxa_df[, c("lon", "lat")],
            proj4string = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
          )

          if (!file.exists(paste0(ps_path, "analyses/ps_", siteoi, "_", taxaoi_short, ".rds"))) {
            # read reflectance data
            files <- list.files(path = paste0(ps_path, siteoi), pattern = ".*_SR_clip.tif$", recursive = T, full.names = T) %>% sort()
            nday <- length(files)
            nloc <- length(plant_taxa_sp)
            ps_mat <- foreach(
              f = 1:nday,
              .packages = c("raster"),
              .combine = "rbind"
            ) %dopar% {
              file <- files[f]
              ps_st <- stack(file)

              trees_sp_reproj <- spTransform(plant_taxa_sp, CRSobj = CRS(proj4string(ps_st)))

              ps_values <- cbind(raster::extract(ps_st, trees_sp_reproj), f, id = 1:nloc)
              print(paste0(f, " out of ", nday))
              ps_values[complete.cases(ps_values), ]
            }

            # read quality assessment data
            # 0 - fully usable data
            # other - potentially problematic/unusable data
            #
            # Full description is in Planet's documentation (Page 91, Section 2. UNUSABLE DATA MASK FILE).
            files <- list.files(path = paste0(ps_path, siteoi), pattern = ".*_udm_clip.tif$", recursive = T, full.names = T) %>% sort()
            nday <- length(files)
            nloc <- length(plant_taxa_sp)
            ps_mask_mat <- foreach(
              f = 1:nday,
              .packages = c("raster"),
              .combine = "rbind"
            ) %dopar% {
              file <- files[f]
              ps_ras <- raster(file)

              trees_sp_reproj <- spTransform(plant_taxa_sp, CRSobj = CRS(proj4string(ps_ras)))

              ps_values <- cbind(qa = raster::extract(ps_ras, trees_sp_reproj), f, id = 1:nloc)

              print(paste0(f, " out of ", nday))
              ps_values[complete.cases(ps_values), ]
            }
            colnames(ps_mask_mat) <- c("qa", "f", "id")
            stopCluster(cl)

            # get corresponding timing from file names
            time_df <- list.files(path = paste0(ps_path, siteoi), pattern = ".*_SR_clip.tif$", recursive = T) %>%
              sort() %>%
              str_split(pattern = "/", simplify = T) %>%
              data.frame() %>%
              dplyr::select(filename = X2) %>%
              rowwise() %>%
              mutate(time = strptime(paste0(str_split(filename, pattern = "_")[[1]][1], str_split(filename, pattern = "_")[[1]][2]), format = "%Y%m%d%H%M%OS")) %>%
              ungroup() %>%
              mutate(f = row_number()) %>%
              dplyr::select(-filename)

            # assign id to each plant
            coord_df <- coordinates(plant_taxa_sp) %>%
              as_tibble() %>%
              mutate(id = row_number())

            # join data
            ps_df <- ps_mat %>%
              as_tibble() %>%
              left_join(time_df, by = "f") %>%
              left_join(coord_df, by = "id") %>%
              mutate(
                red = red * 0.0001, # scaling following Dixon et al's code
                green = green * 0.0001,
                blue = blue * 0.0001,
                nir = nir * 0.0001
              ) %>%
              # mutate(evi=2.5* (nir-red) / (nir + 6*red - 7.5*blue + 1),
              #        gndvi=(nir-green)/(nir+green),
              #        ebi= (red + green + blue) / (green / blue * (red - blue + 1))) %>%
              left_join(ps_mask_mat %>% as_tibble(), by = c("id", "f")) %>%
              dplyr::select(-f)

            # save
            write_rds(ps_df, paste0(ps_path, "analyses/ps_", siteoi, "_", taxaoi_short, ".rds"))
          }
        }
      }
    }
  )

  if (class(iserror) != "try-error") {
    iscomplete <- T
  } else if (class(iserror) == "try-error") { # restart cluster
    iscomplete <- F
    closeAllConnections()
    cl <- makeCluster(50, outfile = "")
    registerDoSNOW(cl)
  }
}
stopCluster(cl)
```

Example time series.
```{r, eval=TRUE}
site_vis <- "DT"
taxa_vis <- "Quercus"
set.seed(1)
read_rds(paste0(ps_path, "analyses/ps_", site_vis, "_", taxa_vis, ".rds")) %>%
  filter(id %in% ((.) %>% pull(id) %>% sample(4))) %>% # four random trees
  filter(qa == 0) %>%
  gather(key = "band", value = "value", -id, -time, -lon, -lat, -qa) %>%
  ggplot() +
  geom_point(aes(x = time, y = value, col = band), alpha = 0.25) +
  facet_wrap(. ~ id, ncol = 1) +
  theme_classic() +
  ggtitle(paste0("Taxa: ", taxa_vis, "; Site: ", site_vis))
```

Compare EVI time series between taxa.
```{r, eval=TRUE}
siteoi <- "SL"
evi_list <- vector(mode = "list")
for (taxaoi_short in taxa_short_list %>% unique()) {
  ps_df <- read_rds(paste0(ps_path, "analyses/ps_", siteoi, "_", taxaoi_short, ".rds"))
  random_id <- ps_df %>%
    pull(id) %>%
    sample(100)
  ps_df_proc <- ps_df %>%
    drop_na() %>%
    filter(id %in% random_id) %>%
    mutate(date = as.Date(time)) %>%
    mutate(
      year = format(time, "%Y") %>% as.integer(),
      doy = format(time, "%j") %>% as.integer(),
      hour = format(strptime(time, "%Y-%m-%d %H:%M:%S"), "%H") %>% as.integer()
    ) %>%
    filter(qa == 0) %>%
    group_by(id, lon, lat, date, year, doy) %>%
    summarise(
      blue = mean(blue),
      green = mean(green),
      red = mean(red),
      nir = mean(nir)
    ) %>%
    ungroup() %>%
    mutate(evi = 2.5 * (nir - red) / (nir + 6 * red - 7.5 * blue + 1)) %>%
    filter(evi > 0, evi <= 1) %>%
    filter(red > 0, green > 0, blue > 0)
  evi_list[[taxaoi_short]] <- ps_df_proc %>% mutate(taxa = taxaoi_short)
}
evi_alltaxa_df <- bind_rows(evi_list)

# ggplot(evi_alltaxa_df %>%
#   filter(taxa %in% c("Poaceae", "Quercus", "Cupressaceae")) %>%
#            filter(year == 2019) %>%
#     filter(doy>=80, doy<200)
#   )+
#   geom_line(aes(x=doy, y=evi, group=id, col=taxa), alpha=0.1)+
#   theme_classic()

ggplot(evi_alltaxa_df %>%
  filter(taxa %in% c("Poaceae", "Quercus", "Cupressaceae")) %>%
  filter(year == 2019) %>%
  filter(doy >= 80, doy < 200) %>%
  group_by(taxa, doy) %>%
  summarise(
    median = median(evi),
    upper = quantile(evi, 0.95),
    lower = quantile(evi, 0.05)
  ) %>%
  ungroup()) +
  geom_line(aes(x = doy, y = median, col = taxa), alpha = 1) +
  geom_ribbon(aes(x = doy, ymin = lower, ymax = upper, fill = taxa), alpha = 0.1) +
  # geom_vline(xintercept = 125)+
  theme_classic() +
  ylab("evi") +
  # facet_wrap(.~taxa, ncol=1)+
  guides(fill = "none")
```

Compare EVI time series between different oak species in Austin.
```{r, eval=TRUE}
siteoi <- "AT"
taxaoi_short <- "Quercus"
ps_df <- read_rds(paste0(ps_path, "analyses/ps_", siteoi, "_", taxaoi_short, ".rds"))
random_id <- ps_df %>%
  pull(id) %>%
  sample(1000)
ps_df_proc <- ps_df %>%
  drop_na() %>%
  filter(id %in% random_id) %>%
  mutate(date = as.Date(time)) %>%
  mutate(
    year = format(time, "%Y") %>% as.integer(),
    doy = format(time, "%j") %>% as.integer(),
    hour = format(strptime(time, "%Y-%m-%d %H:%M:%S"), "%H") %>% as.integer()
  ) %>%
  filter(qa == 0) %>%
  group_by(id, lon, lat, date, year, doy) %>%
  summarise(
    blue = mean(blue),
    green = mean(green),
    red = mean(red),
    nir = mean(nir)
  ) %>%
  ungroup() %>%
  mutate(evi = 2.5 * (nir - red) / (nir + 6 * red - 7.5 * blue + 1)) %>%
  filter(evi > 0, evi <= 1) %>%
  filter(red > 0, green > 0, blue > 0)


plant_taxa_df <- plant_df %>%
  filter(site == siteoi) %>%
  filter(genus == taxaoi_short | family == taxaoi_short) %>%
  mutate(id = row_number()) %>%
  drop_na(lon, lat)

# ggplot(ps_df_proc %>%
#          left_join(plant_taxa_df %>%
#                      dplyr::select(id, species),
#                    by="id") %>%
#            filter(year == 2019)
#   )+
#   geom_line(aes(x=doy, y=evi, group=id, col=species), alpha=0.1)+
#   theme_classic()

ggplot(ps_df_proc %>%
  left_join(plant_taxa_df %>%
    dplyr::select(id, species),
  by = "id"
  ) %>%
  filter(species %in% c("Quercus virginiana", "Quercus fusiformis", "Quercus shumardii")) %>%
  filter(year == 2019) %>%
  group_by(species, doy) %>%
  summarise(
    median = median(evi),
    upper = quantile(evi, 0.95),
    lower = quantile(evi, 0.05)
  ) %>%
  ungroup()) +
  geom_line(aes(x = doy, y = median, col = species), alpha = 1) +
  geom_ribbon(aes(x = doy, ymin = lower, ymax = upper, fill = species), alpha = 0.1) +
  # geom_vline(xintercept = 125)+
  theme_classic() +
  ylab("evi") +
  # facet_wrap(.~taxa, ncol=1)+
  guides(fill = "none")
```