# Infer relationship between leaf-flower lag with climate

Plot correlation between mean annual temperature (MAT) and lag between green-up/down frequency and pollen count.
* A positive lag means leafing phenology leads pollen phenology; a negative lag means leafing phenology lags pollen phenology.
```{r, eval=TRUE}
source("code/clim_plot.R")
p_lag_clim
```
Linear regression to check significance of the correlation. When looking at individual taxa, only 3 were statistically significant.
```{r, eval=TRUE}
source("code/clim_reg.R")
p_slope
```

Linear mixed-effect model with random slope and random intercept to summarize correlation over early and late flowering taxa.
* For early-flowering taxa, pollen phenology usually leads leafing phenology. At warmer places, pollen leads leafing even more. This effect was statistically significant.
* For early-flowering taxa, pollen phenology usually lags leafing phenology. At warmer places, pollen lags green-down even more. This effect was not statistically significant, possibly because there were only three taxa. It was significant for late-flowering elm trees.

```{r,eval=TRUE}
summary(fit_lme)
```
