# Leave-one-out cross validation

We conducted leave-one-out cross validation to test the robustness of the climate-phenology relationship and the effectiveness of using it to infer flowering phenology in new locations. Specifically, we removed a random city from the pollen dataset at a time, matched leafing and pollen phenology in the other cities, and modeled the climate-lag correlation. We predicted the leafing-phenology lag with the linear model and subsequently predicted the flowering phenology from known leafing phenology at the city held for validation. We evaluated the accuracy of our methods by calculating the RMSE between the predicted flowering phenology and standardized pollen count observations at the cities held for validation.

```{r, eval=FALSE}
for (taxaoi in taxa_list) {
  flower_window <- seq(flower_window_df %>% filter(taxa == taxaoi) %>% pull(start),
    flower_window_df %>% filter(taxa == taxaoi) %>% pull(end),
    by = 1
  )

  output_path <- paste0("/raid/users/ysong67/GitHub/phenology/RS4flower/output/data/", taxaoi, "/")

  # read in inferred flowering phenology data
  # flower_doy_df<-read_rds(paste0(output_path,"flowering day of year.rds" ))
  flower_freq_df <- read_rds(paste0(output_path, "flowering frequency.rds"))

  # standardize all data to be between 0 and 1
  flower_freq_df_standard <- flower_freq_df %>%
    mutate(pollen = pollen %>% sqrt()) %>%
    mutate(pollen_sm = pollen_sm %>% sqrt()) %>%
    mutate(pollen_clim = pollen_clim %>% sqrt()) %>%
    group_by(site, sitename, thres) %>%
    mutate(freq = (freq - min(freq, na.rm = T)) / (max(freq, na.rm = T) - min(freq, na.rm = T))) %>%
    mutate(freq_sm = (freq_sm - min(freq_sm, na.rm = T)) / (max(freq_sm, na.rm = T) - min(freq_sm, na.rm = T))) %>%
    mutate(pollen = (pollen - min(pollen, na.rm = T)) / (max(pollen, na.rm = T) - min(pollen, na.rm = T))) %>%
    mutate(pollen_clim = (pollen_clim - min(pollen_sm, na.rm = T)) / (max(pollen_sm, na.rm = T) - min(pollen_sm, na.rm = T))) %>%
    mutate(pollen_sm = (pollen_sm - min(pollen_sm, na.rm = T)) / (max(pollen_sm, na.rm = T) - min(pollen_sm, na.rm = T))) %>%
    mutate(npn = (npn - min(npn, na.rm = T)) / (max(npn, na.rm = T) - min(npn, na.rm = T))) %>%
    mutate(evi = (evi - min(evi, na.rm = T)) / (max(evi, na.rm = T) - min(evi, na.rm = T))) %>%
    mutate(g2r = (g2r - min(g2r, na.rm = T)) / (max(g2r, na.rm = T) - min(g2r, na.rm = T))) %>%
    ungroup()

  # flower_freq_df_check
  flower_freq_df_check <- read_rds(paste0(output_path, "accuracy check.rds"))

  best_thres <- flower_freq_df_check %>%
    group_by(direction, thres) %>%
    summarise(rmse = median(rmse)) %>% # median rmse for each threshold
    arrange(rmse) %>%
    head(1) %>% # keep threshold giving the smallest median rmse
    dplyr::select(direction, thres)

  accuracy_df <- flower_freq_df_check %>%
    right_join(best_thres, by = c("direction", "thres")) %>%
    left_join(meta_df %>% dplyr::select(site, sitename), by = c("region" = "site")) %>%
    mutate(rmse_cv = NA)

  # site_valid_list<-flower_freq_df_check %>% pull(region) %>% unique()
  flower_freq_df_standard_best_list_cv <- vector(mode = "list")
  lag_clim_df_cv_allsite_list <- vector(mode = "list")
  # leave one city out
  for (siteout in site_list) {
    best_thres_cv <- flower_freq_df_check %>%
      filter(region != siteout) %>%
      group_by(direction, thres) %>%
      summarise(rmse = median(rmse)) %>% # median rmse for each threshold
      arrange(rmse) %>%
      head(1) %>% # keep threshold giving the smallest median rmse
      dplyr::select(direction, thres)

    accuracy_df_cv <- flower_freq_df_check %>%
      right_join(best_thres_cv, by = c("direction", "thres")) %>%
      left_join(meta_df %>% dplyr::select(site, sitename), by = c("region" = "site"))

    lag_clim_df_cv <- accuracy_df_cv %>%
      dplyr::select(-rmse, -rmse_ps, -rmse_clim) %>%
      left_join(chelsa_df, by = c("region" = "site")) %>%
      mutate(lag_new = case_when(region != siteout ~ lag)) %>%
      mutate(siteout = siteout) %>%
      mutate(taxa = taxaoi)

    lm.model <- lm(lag_new ~ mat, data = lag_clim_df_cv)

    lag_clim_df_cv <- lag_clim_df_cv %>%
      mutate(lag_fit = predict(lm.model, lag_clim_df_cv))

    lag_clim_df_cv_allsite_list[[siteout]] <- lag_clim_df_cv

    if (siteout %in% (flower_freq_df_check %>% pull(region) %>% unique())) {
      lag_fit <- lag_clim_df_cv %>%
        filter(region == siteout) %>%
        pull(lag_fit) %>%
        round(0)
      # getting time series with new lag
      flower_freq_df_standard_best_cv <- flower_freq_df_standard %>%
        filter(site == siteout) %>%
        filter(
          direction == best_thres_cv$direction,
          thres == best_thres_cv$thres
        ) %>%
        ungroup() %>%
        group_by(site, year)

      if (lag_fit < 0) {
        flower_freq_df_standard_best_cv <- flower_freq_df_standard_best_cv %>%
          mutate(freq_sm = lead(freq_sm, n = -lag_fit)) %>%
          mutate(freq_sm = replace_na(freq_sm, 0))
      } else if (lag_fit == 0) {
        flower_freq_df_standard_best_cv <- flower_freq_df_standard_best_cv %>% mutate(freq_sm = replace_na(freq_sm, 0))
      } else if (lag_fit > 0) {
        flower_freq_df_standard_best_cv <- flower_freq_df_standard_best_cv %>%
          mutate(freq_sm = lag(freq_sm, n = lag_fit)) %>%
          mutate(freq_sm = replace_na(freq_sm, 0))
      }

      pollen_ts_cv <- flower_freq_df_standard_best_cv %>% pull(pollen) # pollen count, for calculating accuracy
      freq_ts_lag_cv <- flower_freq_df_standard_best_cv %>% pull(freq_sm)
      rmse_cv <- sqrt(mean((freq_ts_lag_cv - pollen_ts_cv)^2, na.rm = T)) # rmse between remotely-sensed flowering phenology and pollen count

      accuracy_df[accuracy_df$region == siteout, ]$rmse_cv <- rmse_cv
      flower_freq_df_standard_best_list_cv[[siteout]] <- flower_freq_df_standard_best_cv
    }
    print(paste0(taxaoi, ", ", siteout))
  }
  flower_freq_df_standard_best_cv <- bind_rows(flower_freq_df_standard_best_list_cv)
  lag_clim_df_cv_allsite <- bind_rows(lag_clim_df_cv_allsite_list)

  # make plots
  flower_freq_comp <- ggplot(flower_freq_df_standard_best_cv) +
    geom_point(aes(x = doy, y = npn, col = "flower observation (USA-NPN)"), alpha = 0.5) +
    geom_point(aes(x = doy, y = pollen, col = "pollen count (NAB)")) +
    geom_line(aes(x = doy, y = pollen_clim, col = "pollen count (NAB)"), alpha = 0.5, lwd = 1) +
    geom_point(aes(x = doy, y = evi, col = "EVI (PS)"), alpha = 0.2) +
    geom_line(aes(x = doy, y = freq_sm, col = "flowering frequency"), lwd = 1) +
    theme_classic() +
    facet_wrap(. ~ sitename * year, ncol = 4) +
    scale_color_manual(values = cols) +
    theme(legend.position = "bottom") +
    theme(legend.title = element_blank()) +
    ylab("") +
    ggtitle(paste0("Taxa: ", taxaoi))
  flower_freq_comp

  jpeg(paste0(output_path, "flower frequency compared with other data_cv.jpg"),
    height = 3200, width = 3200, res = 300
  )
  print(flower_freq_comp)
  dev.off()

  flower_freq_corr <- ggplot(flower_freq_df_standard_best_cv %>% mutate(year = as.factor(year))) +
    geom_point(aes(x = freq_sm, y = pollen, group = year, col = year)) +
    geom_smooth(aes(x = freq_sm, y = pollen, group = year, col = year), method = "lm", se = F, lwd = 0.5) +
    geom_smooth(aes(x = freq_sm, y = pollen), method = "lm") +
    theme_classic() +
    facet_wrap(. ~ sitename, ncol = 4) +
    coord_equal() +
    xlim(0, 1) +
    ylim(0, 1) +
    ylab("pollen count^(1/2)") +
    xlab("flowering frequency")
  flower_freq_corr
  jpeg(paste0(output_path, "flower frequency and pollen count correlation_cv.jpg"),
    height = 2400, width = 3200, res = 300
  )
  print(flower_freq_corr)
  dev.off()

  across_site_and_year <- ggplot(flower_freq_df_standard_best_cv %>%
    mutate(year = as.factor(year))) +
    geom_point(aes(x = doy, y = pollen, group = year, col = year)) +
    geom_line(aes(x = doy, y = freq_sm, group = year, col = year)) +
    facet_wrap(. ~ sitename, ncol = 1) +
    theme_classic() +
    ylab("") +
    ggtitle(paste0("Taxa: ", taxaoi))
  across_site_and_year

  jpeg(paste0(output_path, "phenology across site and year_cv.jpg"),
    height = 3200, width = 3200, res = 300
  )
  print(across_site_and_year)
  dev.off()

  write_rds(accuracy_df, paste0(output_path, "accuracy check_cv.rds"))
  write_rds(lag_clim_df_cv_allsite, paste0(output_path, "climate lag.rds"))
}
```

Read in climate-lag relationship and fit LME models.
```{r, eval=TRUE}
lag_clim_df_cv_list <- vector(mode = "list")
for (taxaoi in taxa_list) {
  output_path <- paste0("/raid/users/ysong67/GitHub/phenology/RS4flower/output/data/", taxaoi, "/")
  lag_clim_df_cv_list[[taxaoi]] <- read_rds(paste0(output_path, "climate lag.rds"))
}
lag_clim_df_cv <- bind_rows(lag_clim_df_cv_list) %>%
  mutate(taxa = as.factor(taxa)) %>%
  mutate(taxa = fct_relevel(taxa, levels = taxa_chron)) %>%
  mutate(group = case_when(
    taxa %in% c("Ulmus late", "Poaceae late", "Ambrosia") ~ "late",
    TRUE ~ "early"
  ))

p_val_df_list <- vector(mode = "list")
for (siteoutoi in site_list) {
  p_lag_clim_cv <- ggplot(lag_clim_df_cv %>%
    filter(siteout == siteoutoi) %>%
    filter(!taxa %in% c("Poaceae early", "Poaceae late", "Ambrosia"))) +
    geom_point(aes(x = mat, y = lag_new, col = taxa)) +
    geom_smooth(aes(x = mat, y = lag_new, col = taxa), method = "lm", se = F) +
    ggpubr::stat_cor(
      aes(
        x = mat, y = lag,
        label = paste(..rr.label.., ..p.label.., sep = "*`,`~")
      ),
      p.accuracy = 0.05,
      label.x.npc = "left",
      label.y.npc = "top",
      show.legend = F
    ) +
    theme_classic() +
    facet_wrap(~taxa, scales = "free", ncol = 5) +
    xlab("Mean annual temperature (°C)") +
    ylab("Lag between leafing and pollen phenology (day)") +
    guides(col = F)
  p_lag_clim_cv

  lme.fit <- lme(lag ~ mat, random = ~ 1 + mat | taxa, data = lag_clim_df_cv %>%
    filter(siteout == siteoutoi) %>%
    filter(!taxa %in% c("Poaceae early", "Poaceae late", "Ambrosia")) %>%
    filter(group == "early"), control = lmeControl(opt = "optim", optimMethod = "SANN"))

  p_val_early <- summary(lme.fit)$tTable[2, 5]

  lme.fit <- lme(lag ~ mat, random = ~ 1 + mat | taxa, data = lag_clim_df_cv %>%
    filter(siteout == siteoutoi) %>%
    filter(!taxa %in% c("Poaceae early", "Poaceae late", "Ambrosia")) %>%
    filter(group == "late"), control = lmeControl(opt = "optim", optimMethod = "SANN"))
  p_val_late <- summary(lme.fit)$tTable[2, 5]
  p_val_df_list[[siteoutoi]] <- data.frame(siteout = siteoutoi, early = p_val_early, late = p_val_late)
}
p_val_df <- bind_rows(p_val_df_list) %>%
  mutate(
    early_sig = early < 0.05,
    late_sig = late < 0.05
  )
p_val_df
```


Read in validation results and visualize.
```{r, eval=TRUE}
loocv_df_list <- vector(mode = "list")
for (taxaoi in taxa_list) {
  output_path <- paste0("/raid/users/ysong67/GitHub/phenology/RS4flower/output/data/", taxaoi, "/")
  loocv_df_list[[taxaoi]] <- read_rds(paste0(output_path, "accuracy check_cv.rds")) %>%
    mutate(taxa = taxaoi)
}
loocv_df <- bind_rows(loocv_df_list) %>%
  filter(!taxa %in% c("Poaceae early", "Poaceae late", "Ambrosia")) %>%
  dplyr::select(taxa, city = sitename, rmse_clim, rmse_ps, rmse_cv) %>%
  gather(key = "method", value = "rmse", -taxa, -city) %>%
  mutate(method = case_when(
    method == "rmse_clim" ~ "Climatology",
    method == "rmse_ps" ~ "PS (all cities)",
    method == "rmse_cv" ~ "PS (LOOCV)"
  )) %>%
  mutate(taxa = fct_relevel(taxa, levels = taxa_list)) %>%
  mutate(city = fct_relevel(city, levels = site_lat))


p_rmse_taxa <- ggplot(loocv_df) +
  geom_boxplot(aes(x = taxa, y = rmse, fill = method)) +
  ylab("RMSE") +
  theme_classic() +
  scale_fill_brewer(palette = "Set2")
p_rmse_taxa

p_rmse_city <- ggplot(loocv_df) +
  geom_boxplot(aes(x = city, y = rmse, fill = method)) +
  ylab("RMSE") +
  theme_classic() +
  scale_fill_brewer(palette = "Set2")
p_rmse_city
```

Numerical summaries of RMSE.
```{r, eval=TRUE}
loocv_df %>% 
  group_by(method) %>% 
  summarise(median=median(rmse),
            mean=mean(rmse))

```

