# Leave-one-out cross validation

We conducted leave-one-out cross validation to test the robustness of the climate-phenology relationship and the effectiveness of using it to infer flowering phenology in new locations. Specifically, we removed a random city from the pollen dataset at a time, matched leafing and pollen phenology in the other cities, and modeled the climate-lag correlation. We predicted the leafing-phenology lag with the linear model and subsequently predicted the flowering phenology from known leafing phenology at the city held for validation. We evaluated the accuracy of our methods by calculating the RMSE between the predicted flowering phenology and standardized pollen count observations at the cities held for validation.

```{r, eval=FALSE}
source("code/loocv_tune.R")
```

Read in climate-lag relationship and fit LME models.
```{r, eval=TRUE}
source("code/loocv_lme.R")
```

Read in validation results and visualize.
```{r, eval=TRUE}
source("code/loocv_fit.R")
tb_rmse_out
p_rmse_taxa
# p_rmse_quercus
wtest2
```

