---
title: "RS4flower"
author: "Yiluan Song"
date: "5/12/2022"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE, cache=TRUE)
library(tidyverse)
library(ggrepel)
library(magick)
library(gridExtra)
library(spocc)
library(sf)
library(rnpn)
library(planetR)
unloadNamespace("RJSONIO") # conflict with jsonlite
library(planetR)
library(httr)
library(jsonlite)
library(raster)
library(stringr)
library(pracma)
library(foreach)
library(doSNOW)
library(ptw)
library(EnvCpt)
library(rgdal)
library(mclust)
library(npreg)
library(nlme)
library(readxl)
library(fuzzyjoin)
library(taxize)
library(ptw)
library(segmented)
library(RhpcBLASctl)
library(geosphere)
library(nlme)
library(shiny)
```

* Taxa of interest are based on [Description and allergenic potential of 11 most important pollen taxa in the CUSSC region ranked by percent abundance relative to the sum of all pollen taxa over 31 NAB stations that meet inclusion criteria, 2003–2017] (https://link.springer.com/article/10.1007/s10453-019-09601-2/tables/2)
* Spring- and fall-flowering elms are treated separately as suggested by Allison and Yingxiao.
* I decide to separate grasses into early and late-flowering groups as well because many cities have two pollen peaks.
```{r, eval=TRUE}
taxa_list <- c("Quercus", "Cupressaceae", "Ambrosia", "Morus", "Pinaceae", "Ulmus early", "Ulmus late", "Fraxinus", "Betula", "Poaceae early", "Poaceae late", "Acer", "Populus")
taxa_short_list <- str_split(taxa_list, pattern = " ", simplify = T) [, 1]
```

* Ten major cities in CONUS with pollen count data and street tree inventory.
* Detroit only have location info of oak trees, grasses, and ragweeds.
```{r, eval=TRUE}
site_list <- c("NY", "SJ", "AT", "ST", "HT", "TP", "DT", "DV", "KC", "SL")
sitename_list <- c("New York", "San Jose", "Austin", "Seattle", "Houston", "Tampa", "Detroit", "Denver", "Kansas City", "St. Louis")
```

* Focus on the life cycles from 2018 to 2021, spanning 2017 and 2022.
```{r, eval=TRUE}
year_list <- 2018:2021
```

# 1 Data
## 1.1 NAB data
### 1.1.1 Read and clean data
```{r, eval=FALSE}
nab_path <- "/data/ZHULAB/phenology/nab/2021-10-04/"

# read in station coordinates
station_df <- read_csv("/data/ZHULAB/phenology/nab/NAB stations.csv") %>% # this csv is manually typed
  mutate(id = row_number())

file_list <- list.files(path = nab_path, pattern = ".xlsx", full.names = T)
nab_df_list <- vector(mode = "list", length = length(file_list))
for (i in 1:length(file_list)) {
  file <- file_list[i]

  # read in all data
  dat <- read_excel(
    file,
    col_names = F
  )
  start_n <- which(dat[, 1] == "Date") # excel files have headings of different number of rows. Search for the row(s) that starts with "Date" as the start of data table(s).

  # read in meta data
  if (min(start_n) == 1) { # Some excel files have no meta data before data table
    meta_dat <- NA
    station <- NA
  } else { # If there are meta data before data table, read it as a vector
    meta_dat <- read_excel(
      file,
      col_names = F,
      n_max = start_n[1] - 1
    ) %>% pull()
    station <- meta_dat[2] %>%
      strsplit(split = ": ") %>%
      unlist() %>%
      tail(1) # extract station name from meta data vector
  }

  location <- file %>%
    strsplit(split = c("/")) %>%
    unlist() %>%
    tail(1) %>%
    strsplit(split = " \\(") %>%
    unlist() %>%
    head(1) # city name extracted from file name

  # get coordinates of station
  if (!is.na(station)) { # use both city and station name
    station_info <- data.frame(station = station, location = location) %>%
      stringdist_inner_join(station_df, by = c("location", "station"), max_dist = 20, distance_col = "distance") %>% # join with all entries in station info, measuring dissimilarity in station and location, because their names might be slightly different in both tables
      arrange(station.distance, location.distance) %>%
      head(1) %>% # there might be several close matches. take the closest.
      rename(station = station.y) %>%
      rename(location = location.y) %>%
      dplyr::select(-station.x, -location.x, -station.distance, -location.distance, -distance)
  } else { # use only city name if station name is not available
    station_info <- data.frame(location = location) %>%
      stringdist_inner_join(station_df, by = "location", max_dist = 20, distance_col = "distance") %>%
      arrange(distance) %>%
      head(1) %>%
      rename(location = location.y) %>%
      dplyr::select(-location.x, -distance)
  }

  # Extract pollen and spore data table
  if (length(start_n) == 1) { # only pollen, no spores
    colnum <- ncol(read_excel(
      file,
      skip = start_n[1] - 1
    )) # find number of columns

    pollen_dat <- read_excel(
      file,
      skip = start_n[1] - 1,
      col_types = c(
        "date",
        rep("numeric", colnum - 1)
      )
    )
    if (is.na(pollen_dat[2, 1])) { # meaning finer taxonomic resolutions available
      genus_names <- read_excel(
        file,
        skip = start_n[1] - 1
      ) %>%
        slice(1) %>% # read the first row with finer taxonomic resolution
        gather(key = "old", value = "new") %>% # coarse and fine taxonomy into long format
        mutate(new = case_when(
          is.na(new) ~ old,
          TRUE ~ new
        )) # sometimes only coarse taxonomy is present, then keep that

      pollen_dat <- read_excel(
        file,
        skip = start_n[1] - 1,
        col_types = c(
          "date",
          rep("numeric", colnum - 1)
        ),
        col_names = genus_names$new # change to new colnames with finer taxonomic resolution
      )
    }
    pollen_dat <- pollen_dat %>%
      filter(!is.na(Date)) %>% # filter out rows that are not count data
      gather(key = "taxa", value = "count", -Date) %>% # to long format
      filter(!str_detect(taxa, "Station")) %>% # filter out additional info that are not count data, like "Station Name", "Station Postal Code", "Station State"
      filter(!taxa %in% c("Comment", "WeatherNotes")) %>%
      mutate(group = "pollen") %>%
      rowwise() %>%
      group_by(Date, taxa, group) %>%
      summarize(count = sum(count)) %>%
      ungroup()

    # combine with station info
    nab_df_list[[i]] <- pollen_dat %>%
      cbind(station_info)
  }

  if (length(start_n) == 2) { # both pollen and spores
    colnum_pollen <- ncol(read_excel(
      file,
      skip = start_n[1] - 1,
      n_max = (start_n[2] - 2) - (start_n[1] + 1)
    ))

    pollen_dat <- read_excel(
      file,
      skip = start_n[1] - 1,
      n_max = (start_n[2] - 2) - (start_n[1] + 1),
      col_types = c(
        "date",
        rep("numeric", colnum_pollen - 1)
      )
    )

    if (is.na(pollen_dat[2, 1])) { # meaning finer taxonomic resolutions available
      genus_names <- read_excel(
        file,
        skip = start_n[1] - 1,
        n_max = (start_n[2] - 2) - (start_n[1] + 1)
      ) %>%
        slice(1) %>%
        gather(key = "old", value = "new") %>%
        mutate(new = case_when(
          is.na(new) ~ old,
          TRUE ~ new
        ))

      pollen_dat <- read_excel(
        file,
        skip = start_n[1] - 1,
        n_max = (start_n[2] - 2) - (start_n[1] + 1),
        col_types = c(
          "date",
          rep("numeric", colnum_pollen - 1)
        ),
        col_names = genus_names$new
      )
    }
    pollen_dat <- pollen_dat %>%
      filter(!is.na(Date)) %>%
      gather(key = "taxa", value = "count", -Date) %>%
      filter(!str_detect(taxa, "Station")) %>%
      filter(!taxa %in% c("Comment", "WeatherNotes")) %>%
      mutate(group = "pollen") %>%
      rowwise() %>%
      group_by(Date, taxa, group) %>%
      summarize(count = sum(count)) %>%
      ungroup()

    colname_spore <- ncol(read_excel(
      file,
      skip = start_n[2] - 1
    ))
    spore_dat <- read_excel(
      file,
      skip = start_n[2] - 1,
      col_types = c(
        "date",
        rep("numeric", colname_spore - 1)
      )
    ) %>%
      filter(!is.na(Date)) %>%
      gather(key = "taxa", value = "count", -Date) %>%
      filter(!str_detect(taxa, "Station")) %>%
      filter(!taxa %in% c("Comment", "WeatherNotes")) %>%
      mutate(group = "spore")

    if (nrow(spore_dat) != 0) {
      nab_df_list[[i]] <- bind_rows(pollen_dat, spore_dat) %>%
        cbind(station_info)
    } else {
      nab_df_list[[i]] <- pollen_dat %>%
        cbind(station_info)
    }
  }
  print(i)
}

nab_df <- bind_rows(nab_df_list)

write_rds(nab_df, file = "/data/ZHULAB/phenology/nab/nab_dat.rds")
```

### 1.1.2 Resolve taxonomy
```{r, eval=FALSE}
nab_df <- read_rds("/data/ZHULAB/phenology/nab/nab_dat.rds")

## get all distinct taxa and clean up their names
nab_taxa_df <- nab_df %>%
  distinct(taxa) %>%
  filter(!taxa %in% c("Total Pollen Count", "Total Spore Count")) %>%
  rename(taxa_raw = taxa) %>%
  rowwise() %>%
  mutate(taxa_clean = str_split(taxa_raw, pattern = "/", simplify = T)[1]) %>% # some taxa have alternative names, e.g., Chenopodiaceae/Amaranthaceae
  mutate(taxa_clean = str_split(taxa_clean, pattern = " \\(", simplify = T)[1]) %>% # e.g., Asteraceae (Excluding Ambrosia and Artemisia)
  mutate(taxa_clean = str_replace(taxa_clean, pattern = "-type", "")) %>% # e.g., Leptosphaeria-type
  mutate(taxa_clean = str_replace(taxa_clean, pattern = "Unidentified ", "")) %>% # e.g., Unidentified Fungi
  mutate(taxa_clean = str_replace(taxa_clean, pattern = "Undifferentiated ", "")) %>% # e.g., Undifferentiated Ascospores
  mutate(taxa_clean = str_replace(taxa_clean, pattern = "Other ", "")) %>% # e.g., Other Grass Pollen
  mutate(taxa_clean = str_replace(taxa_clean, pattern = "spores", "mycota")) %>% # e.g., Undifferentiated Ascospores
  mutate(taxa_clean = str_replace(taxa_clean, pattern = " ", "")) %>% # remove space
  mutate(taxa_clean = case_when(
    taxa_clean == "Rusts" ~ "Pucciniales",
    taxa_clean == "Smuts" ~ "Myxomycetes",
    taxa_clean == "Dreshslera" ~ "Helminthosporium",
    (taxa_clean == "GrassPollen" | taxa_clean == "WeedPollen") ~ "Gramineae",
    taxa_clean == "TreePollen" ~ "Tracheophyta",
    taxa_clean == "Pollen" ~ "Viridiplantae",
    TRUE ~ taxa_clean
  )) %>% # manual cleaning
  arrange(taxa_clean)

## Use taxize to resolve taxonomy
# Find sources id
# gnr_datasources()

# match with names in databases
resolve_df <- nab_taxa_df %>%
  pull(taxa_clean) %>%
  gnr_resolve(data_source_ids = c(4, 11), with_context = T, best_match_only = T, fields = "all") %>% # NCBI and GBIF databases
  full_join(nab_taxa_df,
    by = c("user_supplied_name" = "taxa_clean")
  ) %>%
  rename(taxa_clean = user_supplied_name) %>%
  mutate(same = (taxa_clean == matched_name)) # check if all taxa names are valid

# some taxa are incorrectly identified as Metazoa. Resolve again for those.
resolve_df_correct <- resolve_df %>%
  filter(str_detect(classification_path, "Metazoa")) %>%
  pull(taxa_clean) %>%
  gnr_resolve(data_source_ids = c(4, 11), best_match_only = F, fields = "all") %>% # NCBI and GBIF. Keep all matches here, not only the best match.
  filter(!str_detect(classification_path, "Animalia")) %>%
  filter(!str_detect(classification_path, "Metazoa")) %>%
  rename(taxa_clean = user_supplied_name) %>%
  group_by(taxa_clean) %>%
  slice(1) %>%
  ungroup()

# Correct previously resolved taxonomy
resolve_df <- resolve_df %>%
  filter(!str_detect(classification_path, "Metazoa")) %>%
  bind_rows(resolve_df_correct) %>%
  arrange(taxa_clean)

# make sure all taxa are resolved
# resolve_df %>% filter(!same) %>% View()

# get full classification
taxa_id_df <- resolve_df %>%
  dplyr::select(taxa_clean, data_source_id, taxon_id) %>%
  distinct(taxa_clean, .keep_all = T) %>%
  mutate(data_source = case_when(
    data_source_id == 4 ~ "ncbi",
    data_source_id == 11 ~ "gbif"
  ))

taxa_class_df <- vector(mode = "list", length = nrow(taxa_id_df))
for (i in 1:nrow(taxa_id_df)) {
  taxa_class_df[[i]] <-
    classification(taxa_id_df$taxon_id[i], db = taxa_id_df$data_source[i])  [[1]] %>%
    as_tibble() %>%
    filter(rank %in% c("kingdom", "phylum", "class", "order", "family", "genus", "species")) %>%
    mutate(rank = factor(rank, levels = c("kingdom", "phylum", "class", "order", "family", "genus", "species"))) %>%
    dplyr::select(-id) %>%
    spread(key = "rank", value = "name") %>%
    mutate(taxa_clean = taxa_id_df$taxa_clean[i])
}
taxa_class_df <- bind_rows(taxa_class_df)

nab_taxa_df <- nab_taxa_df %>%
  left_join(taxa_class_df, by = "taxa_clean") %>%
  mutate(kingdom = str_replace(kingdom, "Plantae", "Viridiplantae")) %>% # manual correction
  mutate(kingdom = case_when(
    phylum == "Oomycota" ~ "Chromista",
    TRUE ~ kingdom
  ))

write_rds(nab_taxa_df, "/data/ZHULAB/phenology/nab/nab_taxa.rds")
```

Combine nab data and taxa information. Some preprocessing.
```{r, eval=TRUE}
nab_df <- read_rds("/data/ZHULAB/phenology/nab/nab_dat.rds")
nab_taxa_df <- read_rds("/data/ZHULAB/phenology/nab/nab_taxa.rds")

nab_with_taxa_df <- nab_df %>%
  rename(taxa_raw = taxa) %>%
  left_join(nab_taxa_df, by = "taxa_raw") %>%
  rename(taxa = taxa_clean) %>%
  mutate(family = case_when(
    taxa_raw == "Total Pollen Count" ~ "Total",
    TRUE ~ family
  )) %>%
  mutate(genus = case_when(
    taxa_raw == "Total Pollen Count" ~ "Total",
    TRUE ~ genus
  )) %>%
  filter(kingdom == "Viridiplantae" | is.na(kingdom)) %>%
  group_by(Date, lat, lon, station,location, id, family, genus, taxa) %>%
  summarise(count = sum(count)) %>%
  ungroup() %>%
  mutate(date = as.Date(Date)) %>%
  dplyr::select(-Date)
```

Focus on ten cities.
* Exceptions: Denver pollen data are from Colorado Springs; Austin pollen data are from Georgetown; Detroit pollen data are from Sylvania.
```{r, eval=TRUE}
# summarize station info
meta_df <- nab_with_taxa_df %>%
  filter(taxa %in% taxa_short_list) %>% # limit to taxa studied
  drop_na(count) %>%
  group_by(station, location, lat, lon, id) %>%
  summarise(
    mindate = min(date),
    maxdate = max(date),
    n = n()
  ) %>%
  mutate(range = maxdate - mindate) %>%
  ungroup() %>%
  arrange(desc(n)) %>%
  mutate(site = NA)

# match NAB stations with cities studied
meta_df$site[meta_df$id == 26] <- "NY"
meta_df$site[meta_df$id == 7] <- "SJ"
meta_df$site[meta_df$id == 5] <- "AT"
meta_df$site[meta_df$id == 33] <- "ST"
meta_df$site[meta_df$id == 2] <- "HT"
meta_df$site[meta_df$id == 48] <- "TP"
meta_df$site[meta_df$id == 12] <- "DV"
meta_df$site[meta_df$id == 47] <- "DT"
meta_df$site[meta_df$id == 17] <- "KC"
meta_df$site[meta_df$id == 37] <- "SL"
meta_df <- meta_df %>%
  left_join(data.frame(site = site_list, sitename = sitename_list), by = "site")

# make map
p_pollen_map <- ggplot() +
  geom_polygon(data = map_data("state"), aes(x = long, y = lat, group = group), fill = "white") +
  geom_path(data = map_data("state"), aes(x = long, y = lat, group = group), color = "grey50", alpha = 0.5, lwd = 0.2) +
  theme_void() +
  geom_point(data = meta_df, aes(x = lon, y = lat), pch = 10, color = "black", cex = 3) +
  geom_label_repel(data = meta_df %>% filter(site %in% site_list), aes(x = lon, y = lat, label = sitename)) +
  geom_point(data = meta_df %>% filter(site %in% site_list), aes(x = lon, y = lat), pch = 10, color = "red", cex = 3) +
  coord_equal()
p_pollen_map
```

View pollen phenology in study sites.
```{r, eval=TRUE}
nab_with_taxa_df %>%
  left_join(meta_df %>% dplyr::select(id, site, sitename), by = "id") %>%
  filter(site %in% site_list) %>%
  filter(taxa %in% taxa_short_list) %>%
  mutate(doy = format(date, "%j") %>% as.integer()) %>%
  mutate(year = format(date, "%Y") %>% as.integer()) %>%
  # filter(year %in% year_list) %>% 
  group_by(taxa, site) %>% 
  mutate(count=count %>% sqrt()) %>% 
  mutate(count_st=(count-min(count, na.rm = T))/(max(count, na.rm = T)-min(count, na.rm = T))) %>%
  ggplot() +
  geom_point(aes(x = doy, y = count_st, group = taxa, col = taxa), alpha=0.3) +
  facet_grid(cols = vars(taxa), rows = vars(sitename), scales = "free_y") +
  theme_classic()
```

## 1.2 Plant location data
### 1.2.1 Tree inventory
Read in and clean data. Each tree inventory has a different format, so they have to be processed differently.
```{r, eval=FALSE}
cl <- makeCluster(length(site_list), outfile = "")
registerDoSNOW(cl)

trees_df_list <- foreach(
  site = site_list,
  .packages = c("tidyverse", "rgdal")
) %dopar% {
  if (site == "KC" | site == "SL") {
    # Data from urban FIA
    # Download https://experience.arcgis.com/experience/3641cea45d614ab88791aef54f3a1849/page/Urban-Datamart/
    # Manual: https://www.fia.fs.fed.us/library/database-documentation/urban/dbDescription/Urban_FIADB_User_Guides_Database_Description_ver3-0_2021_03_17.pdf

    # tree table
    id_tree_df<-read_csv("/data/ZHULAB/phenology/occurrence/StreetTrees/UrbanFIA/ID_TREE.csv") %>%
    filter(statecd == 29) %>% # Missouri
      filter(countycd %in% case_when(site == "KC" ~ 95, site == "SL" ~ c(189, 510))) %>%
      dplyr::select(plotid, id = cn, spcd, statuscd) %>%
      group_by(id) %>%
      filter(sum(statuscd != 1) == 0) %>% # living trees
      ungroup() %>%
      distinct(id, .keep_all = T) %>% # in case one tree is surveyed repeatedly
      dplyr::select(-statuscd)

    # plot table
    id_plot_df<-read_csv("/data/ZHULAB/phenology/occurrence/StreetTrees/UrbanFIA/ID_PLOT.csv") %>%
    dplyr::select(plotid, lat, lon)

    # species reference table
    ref_sp_df<-read_csv("/data/ZHULAB/phenology/occurrence/StreetTrees/UrbanFIA/REF_SPECIES.csv") %>%
    dplyr::select(spcd, genus, species) %>%
      mutate(species = paste(genus, species))

    # join tables
    trees_df <- id_tree_df %>%
      left_join(id_plot_df, by = "plotid") %>%
      left_join(ref_sp_df, by = "spcd") %>%
      dplyr::select(-plotid, -spcd) %>%
      mutate(site = site)
  }
  if (site == "DT") {
    # Data from Dan Katz
    trees_df <- read_csv("/data/ZHULAB/phenology/occurrence/StreetTrees/Detroit_oak_pheno_obs_spring_2017.csv") %>%
      distinct(id = tree, species = Species, x, y)

    # reproject to WGS84
    pts <- SpatialPoints(trees_df[, c("x", "y")],
      proj4string = CRS("+init=EPSG:3857")
    )
    # +proj=merc +a=6378137 +b=6378137 +lat_ts=0.0 +lon_0=0.0 +x_0=0.0 +y_0=0 +k=1.0 +units=m +nadgrids=@null +wktext +no_defs
    pts_reproj <- spTransform(
      pts,
      CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
    )

    trees_df <- cbind(trees_df %>% dplyr::select(-x, -y), coordinates(pts_reproj)) %>%
      rename(lat = y, lon = x) %>%
      as_tibble() %>%
      mutate(site = site)
  }
  if (site == "DV") {
    # Data from OpenTrees.org 
    trees_df<-read_csv("/data/ZHULAB/phenology/occurrence/StreetTrees/Tree_Inventory_Denver.csv") %>%
    dplyr::select(id = SITE_ID, species = SPECIES_BO, time = INVENTORY_DATE, lat = Y_LAT, lon = X_LONG) %>%
      arrange(desc(time)) %>%
      distinct(id, species, .keep_all = T) %>%
      dplyr::select(-time) %>%
      filter(lon != 0) %>%
      mutate(site = site)
  }
  if (site == "TP") {
    # Data from https://www.opentreemap.org/tampa/map/
    trees_df <- read_csv("/data/ZHULAB/phenology/occurrence/StreetTrees/Tree_Inventory_Tampa.csv") %>%
      dplyr::select(id = `Tree Id`, genus = Genus, species = Species, lat = `Point Y`, lon = `Point X`) %>% # checked that there is no repeated tree id
      mutate(species = paste0(genus, " ", species)) %>%
      dplyr::select(-genus) %>%
      mutate(site = site)
  }
  if (site == "HT") {
    # Data from  https://koordinates.com/layer/25245-houston-texas-street-tree-inventory/data/
    trees_df <- read_csv("/data/ZHULAB/phenology/occurrence/StreetTrees/Tree_Inventory_Houston/houston-texas-street-tree-inventory.csv") %>%
      dplyr::select(species_common = Species, X = Shape_X, Y = Shape_Y) %>%
      rowwise() %>%
      mutate(common = case_when(
        str_detect(species_common, ", spp.") ~ str_replace(species_common, ", spp.", ""), # for coarse common names, e.g., "Oak, spp."
        (str_detect(species_common, ", ") & !str_detect(species_common, ", spp.")) ~ paste0(str_split(species_common, ", ")[[1]][2], " ", str_split(species_common, ", ")[[1]][1]), # reverse order of common name, e.g., "Oak, Water"
        TRUE ~ species_common
      )) %>%
      ungroup() %>%
      distinct(X, Y, .keep_all = T) %>% # in case same tree is sampled repeatedly
      mutate(id = row_number())

    # reproject to WGS84
    pts <- SpatialPoints(trees_df[, c("X", "Y")],
      proj4string = CRS("+proj=merc +a=6378137 +b=6378137 +lat_ts=0 +lon_0=0 +x_0=0 +y_0=0 +k=1 +units=m +nadgrids=@null +wktext +no_defs")
    )
    pts_reproj <- spTransform(
      pts,
      CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
    )

    trees_df <- cbind(trees_df %>% dplyr::select(-X, -Y), coordinates(pts_reproj)) %>%
      rename(lat = Y, lon = X)

    # Use taxize to match common name with scientific name
    if (!file.exists("/data/ZHULAB/phenology/occurrence/StreetTrees/Tree_Inventory_Houston/species_reference.csv")) {
      id2comm <- function(pageid) {
        common_matches <- eol_pages(taxonconceptID = pageid, common_names = TRUE)$vernacular
        if (is.null(common_matches)) {
          common_match <- ""
        } else {
          common_matches <- common_matches %>% filter(language == "en")
          if (nrow(common_matches) > 0) {
            common_match <- common_matches %>%
              pull(vernacularname) %>%
              table() %>%
              sort(decreasing = TRUE) %>%
              head(1) %>%
              names()
          }
        }
        return(common_match)
      }

      comm2sci_new <- function(common) {
        species <- comm2sci(common, simplify = T)[[1]]
        if (length(species) == 0) {
          res <- comm2sci(common, db = "eol", simplify = F)[[1]]
          if (length(res) == 0) {
            species <- NA
          } else {
            species <- res %>%
              as_tibble() %>%
              rowwise() %>%
              mutate(common_match = id2comm(pageid)) %>%
              ungroup() %>%
              stringdist_inner_join(data.frame(common_name = common), by = c("common_match" = "common_name"), max_dist = 20, distance_col = "distance") %>%
              arrange(distance, pageid) %>%
              head(1) %>%
              filter(common_match != "") %>%
              pull(name)
            if (length(species) == 0) {
              species <- NA
            }
          }
          print(paste(common, species))
        }

        if (is.na(species)) {
          species <- readline(prompt = paste0(species_df$common[i], ". Manually enter scientific name: "))
        }

        return(species)
      }

      species_df <- trees_df %>%
        distinct(common) %>%
        rowwise() %>%
        mutate(species = comm2sci_new(common))
      write_csv(species_df, "/data/ZHULAB/phenology/occurrence/StreetTrees/Tree_Inventory_Houston/species_reference.csv")
    } else {
      species_df <- read_csv("/data/ZHULAB/phenology/StreetTrees/Tree_Inventory_Houston/species_reference.csv")
    }

    trees_df <- trees_df %>%
      left_join(species_df, by = "common") %>%
      dplyr::select(-species_common, -common) %>%
      mutate(site = site)
  }
  if (site == "NY") {
    trees_df <- read_csv("/data/ZHULAB/phenology/occurrence/StreetTrees/Tree_Inventory_NewYork.csv") %>%
      dplyr::select(id = tree_id, species = spc_latin, lat = latitude, lon = longitude) %>% # checked that there is no repeated tree id
      mutate(site = site)
  }
  if (site == "AT") {
    trees_df <- read_csv("/data/ZHULAB/phenology/occurrence/StreetTrees/Tree_Inventory_Austin/Tree_Inventory_Austin.csv") %>%
      dplyr::select(id = OBJECTID, species = SPECIES, coordinates = the_geom) %>% # checked that there is no repeated tree id
      mutate(coordinates = str_replace(coordinates, "POINT \\(", "")) %>%
      mutate(coordinates = str_replace(coordinates, "\\)", "")) %>%
      rowwise() %>%
      mutate(
        lon = str_split(coordinates, pattern = " ", simplify = T)[1],
        lat = str_split(coordinates, pattern = " ", simplify = T)[2]
      ) %>%
      ungroup() %>%
      mutate(
        lon = as.numeric(lon),
        lat = as.numeric(lat)
      ) %>%
      dplyr::select(-coordinates) %>%
      mutate(site = site)
  }
  if (site == "SJ") {
    trees_df <- read_csv("/data/ZHULAB/phenology/occurrence/StreetTrees/Tree_Inventory_SanJose/Tree_Inventory_SanJose.csv") %>%
      dplyr::select(id = OBJECTID, species = NAMESCIENTIFIC, Y = Y, X = X) # checked that there is no repeated tree id

    # shape <- readOGR(dsn = "/data/ZHULAB/phenology/occurrence/StreetTrees/Tree_Inventory_SanJose/Street_Tree.shp")
    # proj4string(shape)
    projection_sj <- "+proj=lcc +lat_0=36.5 +lon_0=-120.5 +lat_1=38.4333333333333 +lat_2=37.0666666666667 +x_0=2000000.0001016 +y_0=500000.0001016 +datum=NAD83 +units=us-ft +no_defs"
    pts <- SpatialPoints(trees_df[, c("X", "Y")],
      proj4string = CRS(projection_sj)
    )
    pts_reproj <- spTransform(
      pts,
      CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
    )

    trees_df <- cbind(trees_df %>% dplyr::select(-X, -Y), coordinates(pts_reproj)) %>%
      rename(lat = Y, lon = X) %>%
      mutate(site = site)
  }

  if (site == "ST") {
    trees_df <- read_csv("/data/ZHULAB/phenology/occurrence/StreetTrees/Tree_Inventory_Seattle.csv") %>%
      dplyr::select(id = OBJECTID, species = SCIENTIFIC_NAME, lat = Y, lon = X) %>% # checked that there is no repeated tree id
      mutate(site = site)
  }
  trees_df
}

stopCluster(cl)

trees_df <- bind_rows(trees_df_list) %>%
  rowwise() %>%
  mutate(genus = str_split(species, pattern = " ", simplify = T)[1]) %>% # get genus name from species name
  ungroup() %>%
  distinct(id, site, .keep_all = T) %>% # in case one tree is surveyed repeatedly
  mutate(genus_id = as.integer(as.factor(genus)))
```

Find family names from genus names. This step needs supervision.
```{r, eval=FALSE}
genus_to_family<-trees_df %>%
  distinct(genus) %>% 
  drop_na() %>% 
  mutate(family=tax_name(sci = genus, get = "family", messages = F, db="ncbi")$family)
write_rds(genus_to_family, "/data/ZHULAB/phenology/occurrence/genus_to_family.rds")
```

Join with inventory data and save.
```{r, eval=FALSE}
genus_to_family<-read_rds("/data/ZHULAB/phenology/occurrence/genus_to_family.rds")

trees_df<-trees_df %>% 
  left_join(genus_to_family, by="genus")
write_rds(trees_df, "/data/ZHULAB/phenology/occurrence/street_trees.rds")
```

### 1.2.2 Grass patch
```{r, eval=FALSE}
# Land cover data manually downloaded from Sentinel-2 Land Use/ Land Cover Downloader
# https://www.arcgis.com/apps/instant/media/index.html?appid=fc92d38533d440078f17678ebc20e8e2
grass_path <- "/data/ZHULAB/phenology/occurrence/LULC/"
trees_df <- read_rds("/data/ZHULAB/phenology/occurrence/street_trees.rds")

cl <- makeCluster(length(site_list), outfile = "")
registerDoSNOW(cl)

grass_df_city_list <- foreach(
  siteoi = site_list,
  .packages = c("tidyverse", "raster")
) %dopar% {

  # get extent from tree inventory
  trees_df_city <- trees_df %>%
    filter(site == siteoi) %>%
    drop_na(lon, lat)
  bbox <- extent(min(trees_df_city$lon), max(trees_df_city$lon), min(trees_df_city$lat), max(trees_df_city$lat))

  grass_df_city_year_list <- vector(mode = "list")
  for (yearoi in year_list) {
    # get file name(s) for each site and year
    files <- list.files(paste0(grass_path, siteoi), pattern = paste0(yearoi, "0101-"), full.names = T)
    
    # read raster(s)
    ras <- raster(files)
    
    # crop to city
    bbox_sp <- as(bbox, "SpatialPolygons")
    projection(bbox_sp) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
    bbox_reproj <- spTransform(bbox_sp, proj4string(ras))
    ras_cr <- crop(ras, bbox_reproj)

    # get coordinates of grass
    grass_df_year <- as.data.frame(ras_cr, xy = T) %>%
      `colnames<-`(c("x", "y", "class")) %>%
      filter(class == 11) %>%
      dplyr::select(-class) %>%
      mutate(year = yearoi)
    grass_df_city_year_list[[yearoi %>% as.character()]] <- grass_df_year
    print(paste0(siteoi, yearoi))
  }
  grass_df_city <- bind_rows(grass_df_city_year_list) %>%
    mutate(grass = 1) %>% # choose pixels that are grass in all years
    spread(key = "year", value = "grass") %>%
    drop_na() %>%
    dplyr::select(x, y) %>%
    sample_n(min(10000, nrow(.))) # subset when there are too many grass pixels

  # reproject
  grass_sp_reproj <- SpatialPoints(grass_df_city[, c("x", "y")],
    proj4string = CRS(proj4string(ras))
  )
  grass_sp <- spTransform(
    grass_sp_reproj,
    CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
  )

  # get coordinates
  grass_df_city <- coordinates(grass_sp) %>%
    as_tibble() %>%
    `colnames<-`(c("lon", "lat")) %>%
    mutate(site = siteoi) %>%
    mutate(id = row_number()) %>%
    mutate(family = "Poaceae") %>%
    mutate(genus = "Unknown") %>%
    mutate(genus_id = 999)
  grass_df_city
}
stopCluster(cl)

grass_df <- bind_rows(grass_df_city_list)
write_rds(grass_df, "/data/ZHULAB/phenology/occurrence/grass.rds")
```

### 1.2.3 Ragweed observation
Get ragweed occurrence information from GBIF.
```{r, eval=FALSE}
trees_df <- read_rds("/data/ZHULAB/phenology/occurrence/street_trees.rds")

cl <- makeCluster(length(site_list), outfile = "")
registerDoSNOW(cl)

ragweed_df_city_list <- foreach(
  siteoi = site_list,
  .packages = c("tidyverse", "raster", "sf", "spocc")
) %dopar% {
  # get extent
  trees_df_city <- trees_df %>%
    filter(site == siteoi) %>%
    drop_na(lon, lat)
  bbox <- extent(min(trees_df_city$lon), max(trees_df_city$lon), min(trees_df_city$lat), max(trees_df_city$lat))
  # make it a polygon
  bbox_sp <- as(bbox, "SpatialPolygons")
  projection(bbox_sp) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"

  # get gbif data
  res <- occ(
    query = "Ambrosia", from = "gbif", has_coords = TRUE, limit = 1e6,
    geometry = st_bbox(bbox_sp),
    date = c(as.Date("2018-01-01"), as.Date("2021-12-31")),
    gbifopts = list(
      hasGeospatialIssue = FALSE
    )
  )

  # get coordinates
  ragweed_df_city <- res$gbif$data[[1]] %>%
    dplyr::select(lon = longitude, lat = latitude, species) %>%
    mutate(site = siteoi) %>%
    mutate(family = "Asteraceae") %>%
    mutate(genus = "Ambrosia") %>%
    mutate(genus_id = 998)
  ragweed_df_city
}
stopCluster(cl)

ragweed_df <- bind_rows(ragweed_df_city_list)
write_rds(ragweed_df, "/data/ZHULAB/phenology/occurrence/ragweed.rds")
```

### 1.2.4 View plant occurrence data

Put tree, grass, and ragweed data in one data frame.
```{r, eval=TRUE}
trees_df <- read_rds("/data/ZHULAB/phenology/occurrence/street_trees.rds")
grass_df <- read_rds("/data/ZHULAB/phenology/occurrence/grass.rds")
ragweed_df <- read_rds("/data/ZHULAB/phenology/occurrence/ragweed.rds")
plant_df <- bind_rows(trees_df, grass_df, ragweed_df)
```

Map relative position of tree inventory and nab station.
```{r, eval=TRUE}
ggplot() +
  geom_polygon(data = map_data("state"), aes(x = long, y = lat, group = group), fill = "white") +
  geom_path(data = map_data("state"), aes(x = long, y = lat, group = group), color = "grey50", alpha = 0.5, lwd = 0.2) +
  theme_void() +
  geom_point(
    data = plant_df %>%
      filter(genus %in% taxa_short_list | family %in% taxa_short_list) %>%
      mutate(taxa = case_when(
        family %in% c("Poaceae", "Cupressaceae", "Pinaceae") ~ family,
        TRUE ~ genus
      )) %>%
      # filter(!taxa%in% c("Poaceae", "Ambrosia")) %>%
      group_by(site) %>%
      sample_n(100) %>%
      ungroup(),
    aes(x = lon, y = lat), col = "dark blue", alpha = 0.5
  ) +
  geom_point(data = meta_df %>% filter(site %in% site_list), aes(x = lon, y = lat), cex = 3, pch = 10, col = "red", alpha = 1) +
  # geom_label_repel(data=meta_df %>% filter(site %in% site_list), aes(x=lon, y=lat, label = sitename))+
  coord_equal()
```

Calculate distance from plants to NAB stations in the unit of km.
```{r, eval=FALSE}
distance_df<-plant_df %>%
  filter(genus %in% taxa_short_list | family %in% taxa_short_list) %>% 
  left_join(meta_df %>% dplyr::select (site, sitelon=lon, sitelat=lat, sitename) %>% drop_na(), by="site") %>% 
  rowwise() %>% 
  mutate(distance=distm(c(lon, lat), c(sitelon, sitelat), fun = distHaversine) %>% as.numeric() %>% `/`(1000)) %>% # distance in the unit of km
  ungroup() %>% 
  group_by(site, sitename) %>% 
  summarise(mindist=min(distance),
            maxdist=max(distance),
            meandist=mean(distance))
write_rds(distance_df, "/data/ZHULAB/phenology/occurrence/distance_from_plants_to_nab_stations.rds")
```

```{r, eval=TRUE}
read_rds("/data/ZHULAB/phenology/occurrence/distance_from_plants_to_nab_stations.rds")
```

Map plant occurrence in each city.
```{r, eval=TRUE}
ggplot() +
  theme_classic() +
  geom_point(
    data = plant_df %>%
      filter(genus %in% taxa_short_list | family %in% taxa_short_list) %>%
      mutate(taxa = case_when(
        family %in% c("Poaceae", "Cupressaceae", "Pinaceae") ~ family,
        TRUE ~ genus
      )) %>%
      # filter(!taxa%in% c("Poaceae", "Ambrosia")) %>%
      group_by(taxa, site) %>%
      sample_n(min(100, n())) %>%
      ungroup() %>%
      left_join(meta_df %>% dplyr::select(site, sitename), by = "site"),
    aes(x = lon, y = lat, col = taxa), alpha = 0.3
  ) +
  facet_wrap(. ~ sitename, scales = "free")
```

## 1.3 NPN data
```{r, eval=FALSE}
phenophases <- npn_phenophases()
species_list <- npn_species()

# download all NPN data for taxa studied
npn_path <- "/data/ZHULAB/phenology/NPN/"

cl <- makeCluster(length(site_list), outfile = "")
registerDoSNOW(cl)

npn_df_list <- vector(mode = "list")
for (taxaoi_short in taxa_short_list %>% unique()) {
  if (!file.exists(paste0(npn_path, taxaoi_short, ".rds"))) {
    spid <- species_list %>%
      filter(genus == taxaoi_short | family_name == taxaoi_short) %>%
      pull(species_id)

    npn_data <- npn_download_status_data(request_source = "YS", years = c(2000:2022), species_id = spid)

    write_rds(npn_data, paste0(npn_path, taxaoi_short, ".rds"))
  } else {
    npn_data <- read_rds(paste0(npn_path, taxaoi_short, ".rds"))
  }

  npn_taxa_df_list <- foreach(
    siteoi = site_list,
    .packages = c("tidyverse", "geosphere")
  ) %dopar% {
    lat_oi <- meta_df %>%
      filter(site == siteoi) %>%
      pull(lat) %>%
      mean()
    lon_oi <- meta_df %>%
      filter(site == siteoi) %>%
      pull(lon) %>%
      mean()

    npn_flower_df <- npn_data %>%
      filter(phenophase_status == 1) %>%
      filter(phenophase_description %in% c("Full pollen release (conifers)", "Pollen release (conifers)", "Pollen cones (conifers)", "Open pollen cones (conifers)", "Full flowering (50%)", "Flowers or flower buds", "Pollen release (flowers)"))

    if (nrow(npn_flower_df > 0)) {
      npn_flower_df <- npn_flower_df %>%
        rowwise() %>%
        mutate(distance = distm(c(longitude, latitude), c(lon_oi, lat_oi), fun = distHaversine) %>% as.numeric()) %>% # distance in the unit of m
        ungroup() %>%
        filter(distance <= 500000) %>% # within 500 km of the NAB station
        dplyr::select(date = observation_date, lon = longitude, lat = latitude, doy = day_of_year) %>%
        mutate(date = as.Date(date)) %>%
        mutate(site = siteoi)
    } else {
      npn_flower_df <- data.frame(lon = numeric(0), lat = numeric(0), doy = integer(0), date = character(0), site = character(0)) %>% mutate(date = as.Date(date))
    }

    npn_count_df <- npn_flower_df %>%
      group_by(site, date) %>%
      summarise(count = n()) %>%
      ungroup()

    print(paste0(taxaoi_short, ", ", siteoi))
    npn_count_df
  }

  npn_df_list[[taxaoi_short]] <- bind_rows(npn_taxa_df_list) %>%
    mutate(taxa = taxaoi_short)
}
stopCluster(cl)
npn_df <- bind_rows(npn_df_list)
write_rds(npn_df, paste0(npn_path, "npn_dat.rds"))
```

View flowering phenology (from NPN data) in study sites.
```{r, eval=TRUE}
npn_path <- "/data/ZHULAB/phenology/NPN/"
npn_df_all<-read_rds(paste0(npn_path, "npn_dat.rds"))
npn_df_all %>%
  left_join(meta_df %>% dplyr::select(id, site, sitename), by = "site") %>%
  filter(site %in% site_list) %>%
  filter(taxa %in% taxa_short_list) %>%
  mutate(doy = format(date, "%j") %>% as.integer()) %>%
  mutate(year = format(date, "%Y") %>% as.integer()) %>%
  # filter(year %in% year_list) %>% 
  group_by(taxa, site) %>% 
  mutate(count_st=(count-min(count, na.rm = T))/(max(count, na.rm = T)-min(count, na.rm = T))) %>%
  ggplot() +
  geom_point(aes(x = doy, y = count_st, group = taxa, col = taxa), alpha=0.3) +
  facet_grid(cols = vars(taxa), rows = vars(sitename), scales = "free_y") +
  theme_classic()
```

## 1.4 CHELSA data
```{r, eval=TRUE}
# Downloaded from: https://chelsa-climate.org/bioclim/
# Documentation: https://chelsa-climate.org/wp-admin/download-page/CHELSA_tech_specification_V2.pdf

chelsa_path <- "/data/ZHULAB/phenology/CHELSA/"
# read in raster
tmean_ras <- raster(paste0(chelsa_path, "bio1.tif"))
ppt_ras <- raster(paste0(chelsa_path, "bio12.tif"))
vpdmax_ras <- raster(paste0(chelsa_path, "vpd_max.tif"))

# sites as points
meta_sf <- meta_df %>%
  drop_na(site) %>%
  dplyr::select(site, lon, lat) %>%
  st_as_sf(
    coords = c("lon", "lat"),
    crs = 4326
  )

# extract chelsa data at points
chelsa_df <- meta_sf %>%
  mutate(
    mat = raster::extract(tmean_ras, meta_sf),
    tap = raster::extract(ppt_ras, meta_sf),
    vpd = raster::extract(vpdmax_ras, meta_sf)
  ) %>%
  as_tibble() %>%
  dplyr::select(-geometry)
```

## 1.5 PlanetScope data
### 1.5.1 Order

```{r, eval=FALSE, echo=FALSE}
# Modify a function to record order name
planet_order_download_new <- function(order_id, order_name, api_key, order_num, overwrite_opt = FALSE) {
  # GET order for download
  # If you lose the order_id, don't redo the request, log onto planet and find it in the orders menu
  # order_id for example SMV2 order: "dab92990-ce3a-456c-8ad6-ca0c569b4a1a"
  url2 <- paste0("https://api.planet.com/compute/ops/orders/v2/", order_id)

  get_order <- httr::GET(
    url = url2,
    username = api_key
  )
  # Download links are in here, under _links>results>location
  get_content <- httr::content(get_order)
  # When state = 'success', ready for download

  # check if order is ready
  while (get_content$state != "success") {
    print("Order still being proccessed, trying again in 60 seconds...")
    print(get_content$state)
    Sys.sleep(60)
    get_order <- httr::GET(url = url2, username = api_key)
    get_content <- httr::content(get_order)
  }

  ## Time to download!
  print("Starting download")

  # First create download folder:
  dir.create(order_name, showWarnings = F)

  # Download each item in order
  for (i in 1:length(get_content$`_links`$results)) {
    print(paste0("Order ", order_num, ", Download: ", round(100 * (
      i / length(get_content$`_links`$results)
    ), 2), "%"))
    # find item names in order contents
    name <- get_content$`_links`$results[[i]]$name
    findslash <- gregexpr("/", name)
    startchar <- findslash[[1]][length(findslash[[1]])] + 1
    filename <- substr(name, startchar, nchar(name))

    download_url <- get_content$`_links`$results[[i]]$location

    try({
      httr::RETRY(
        "GET",
        url = download_url,
        username = api_key,
        write_disk(
          path = paste(order_name, filename, sep = "/"),
          overwrite = overwrite_opt
        )
      )
    })
  }

  print(paste0("Download complete"))
  print(paste0("Items located in ", order_name))
}
```

User settings.
```{r, eval=TRUE}
# Set API
# api_key = "REMOVED" #ysong67
api_key <- "REMOVED" # xcui12

# Metadata filters
cloud_lim <- 0.99 # percent from 0-1
item_name <- "PSScene4Band"
# PSOrthoTile, PSScene3Band, PSScene4Band, Sentinel2L1C
# (see https://developers.planet.com/docs/data/items-assets/)
product <- "analytic_sr"
# analytic_b1, analytic_b2
# (see https://developers.planet.com/docs/data/items-assets/)

ps_path <- "/data/ZHULAB/phenology/PlanetScope/"
```

Order data. Orders may fail. May have to order again in that case.
```{r, eval=FALSE}
for (siteoi in site_list) {
  ps_path_site <- paste0(ps_path, siteoi, "/")
  dir.create(ps_path_site, recursive = T)
  dir.create(paste0(ps_path_site, "orders/"), recursive = T)

  # Set AOI (many ways to set this!) ultimately just need an extent()
  plant_df_site <- plant_df %>%
    filter(site == siteoi) %>%
    drop_na(lon, lat)
  bbox <- extent(min(plant_df_site$lon), max(plant_df_site$lon), min(plant_df_site$lat), max(plant_df_site$lat))

  for (year_download in 2017:2022) {
    order_df <- data.frame(year = integer(0), month = integer(0), id = character(0))
    if (year_download == 2022) {
      month_end <- 4
    } else {
      month_end <- 12
    }
    for (month_download in 1:month_end) {
      # Date range of interest
      start_year <- year_download
      end_year <- year_download
      date_start <- lubridate::floor_date(as.Date(paste0(year_download, "-", str_pad(month_download, 2, pad = "0"), "-01")), unit = "month")
      date_end <- lubridate::ceiling_date(as.Date(paste0(year_download, "-", str_pad(month_download, 2, pad = "0"), "-01")), unit = "month") - 1
      start_doy <- as.numeric(format(date_start, "%j"))
      end_doy <- as.numeric(format(date_end, "%j"))

      # Create order name
      order_name <- paste(siteoi, item_name, product, start_year, end_year, start_doy, end_doy, sep = "_")

      # Planet Orders API
      order_id <- planetR:::planet_order_request(
        api_key = api_key,
        bbox = bbox,
        date_start = date_start,
        date_end = date_end,
        start_doy = start_doy,
        end_doy = end_doy,
        cloud_lim = cloud_lim,
        item_name = item_name,
        product = product,
        order_name = order_name
      )

      # when there are too many items in an order, order id will not be returned. split order into a few groups in that case.
      if (!is.null(order_id)) {
        # store order name
        order_df <- order_df %>%
          bind_rows(data.frame(year = year_download, month = month_download, id = order_id))
      } else {
        item_num <- nrow(planet_search(
          bbox, start_doy, end_doy, date_end,
          date_start, cloud_lim, item_name, api_key
        ))
        group_num <- ceiling(item_num / 500)
        split_num <- ceiling(30 / group_num)
        doy_group <- split(seq(start_doy, end_doy), floor(seq(1:(end_doy - start_doy + 1)) / split_num))
        for (g in 1:length(doy_group)) {
          order_id <- planetR:::planet_order_request(
            api_key = api_key,
            bbox = bbox,
            date_start = date_start,
            date_end = date_end,
            start_doy = min(doy_group[[g]]),
            end_doy = max(doy_group[[g]]),
            cloud_lim = cloud_lim,
            item_name = item_name,
            product = product,
            order_name = order_name
          )
          if (!is.null(order_id)) {
            order_df <- order_df %>%
              bind_rows(data.frame(year = year_download, month = month_download, id = order_id))
          }
        }
      }

      print(paste(year_download, ", ", month_download))
    }
    dir.create(paste0(ps_path_site, "orders/"))
    write_rds(order_df, paste0(ps_path_site, "orders/", "order_", year_download, ".rds"))
  }
}
```

### 1.5.2 Download
```{r, eval=FALSE}
for (siteoi in site_list) {
  ps_path_site <- paste0(ps_path, siteoi, "/")
  for (year_download in 2017:2022) {
    order_df <- read_rds(paste0(ps_path_site, "orders/", "order_", year_download, ".rds"))
    cl <- makeCluster(nrow(order_df), outfile = "")
    registerDoSNOW(cl)
    foreach(
      i = 1:nrow(order_df),
      .packages = c("stringr", "planetR", "httr")
    ) %dopar% {

      # Get order id
      month_download <- order_df$month[i]
      order_id <- order_df$id[i]

      # Date range of interest
      start_year <- year_download
      end_year <- year_download
      date_start <- lubridate::floor_date(as.Date(paste0(year_download, "-", str_pad(month_download, 2, pad = "0"), "-01")), unit = "month")
      date_end <- lubridate::ceiling_date(as.Date(paste0(year_download, "-", str_pad(month_download, 2, pad = "0"), "-01")), unit = "month") - 1
      start_doy <- as.numeric(format(date_start, "%j"))
      end_doy <- as.numeric(format(date_end, "%j"))

      # Set/Create Export Folder
      order_name <- paste(siteoi, item_name, product, start_year, end_year, start_doy, end_doy, sep = "_")
      exportfolder <- paste0(ps_path_site, order_name)
      dir.create(exportfolder, recursive = T, showWarnings = F)

      # Download
      Sys.sleep(i * 0.5) # Otherwise sending request to API at the same time may cause error
      planet_order_download_new(order_id, exportfolder, api_key = api_key, order_num = i, overwrite_opt = FALSE)

      print(paste(year_download, ", ", month_download))
    }

    stopCluster(cl)
  }
}
```

### 1.5.3 Retrieve time series at plant locations
```{r, eval=FALSE}
cl <- makeCluster(50, outfile = "")
registerDoSNOW(cl)

iscomplete <- F
while (!iscomplete) { # restart when there is error, usually because of cluster connection issues
  iserror <- try(
    for (taxaoi in taxa_list) {
      taxaoi_short <- str_split(taxaoi, " ", simplify = T)[1]
      for (s in 1:length(site_list)) {
        # get plant locations
        siteoi <- site_list[s]
        plant_taxa_df <- plant_df %>%
          filter(site == siteoi) %>%
          filter(genus == taxaoi_short | family == taxaoi_short) %>%
          mutate(id = row_number()) %>%
          drop_na(lon, lat)

        # skip when there are too few plants
        if (taxaoi_short %in% c("Ambrosia", "Poaceae")) {
          min_sample_size <- 1
        } else {
          min_sample_size <- 1
        }

        if (nrow(plant_taxa_df) >= min_sample_size) {
          # plants as points
          plant_taxa_sp <- SpatialPoints(plant_taxa_df[, c("lon", "lat")],
            proj4string = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
          )

          if (!file.exists(paste0(ps_path, "analyses/ps_", siteoi, "_", taxaoi_short, ".rds"))) {
            # read reflectance data
            files <- list.files(path = paste0(ps_path, siteoi), pattern = ".*_SR_clip.tif$", recursive = T, full.names = T) %>% sort()
            nday <- length(files)
            nloc <- length(plant_taxa_sp)
            ps_mat <- foreach(
              f = 1:nday,
              .packages = c("raster"),
              .combine = "rbind"
            ) %dopar% {
              file <- files[f]
              ps_st <- stack(file)

              trees_sp_reproj <- spTransform(plant_taxa_sp, CRSobj = CRS(proj4string(ps_st)))

              ps_values <- cbind(raster::extract(ps_st, trees_sp_reproj), f, id = 1:nloc)
              print(paste0(f, " out of ", nday))
              ps_values[complete.cases(ps_values), ]
            }

            # read quality assessment data
            # 0 - fully usable data
            # other - potentially problematic/unusable data
            #
            # Full description is in Planet's documentation (Page 91, Section 2. UNUSABLE DATA MASK FILE).
            files <- list.files(path = paste0(ps_path, siteoi), pattern = ".*_udm_clip.tif$", recursive = T, full.names = T) %>% sort()
            nday <- length(files)
            nloc <- length(plant_taxa_sp)
            ps_mask_mat <- foreach(
              f = 1:nday,
              .packages = c("raster"),
              .combine = "rbind"
            ) %dopar% {
              file <- files[f]
              ps_ras <- raster(file)

              trees_sp_reproj <- spTransform(plant_taxa_sp, CRSobj = CRS(proj4string(ps_ras)))

              ps_values <- cbind(qa = raster::extract(ps_ras, trees_sp_reproj), f, id = 1:nloc)

              print(paste0(f, " out of ", nday))
              ps_values[complete.cases(ps_values), ]
            }
            colnames(ps_mask_mat) <- c("qa", "f", "id")
            stopCluster(cl)

            # get corresponding timing from file names
            time_df <- list.files(path = paste0(ps_path, siteoi), pattern = ".*_SR_clip.tif$", recursive = T) %>%
              sort() %>%
              str_split(pattern = "/", simplify = T) %>%
              data.frame() %>%
              dplyr::select(filename = X2) %>%
              rowwise() %>%
              mutate(time = strptime(paste0(str_split(filename, pattern = "_")[[1]][1], str_split(filename, pattern = "_")[[1]][2]), format = "%Y%m%d%H%M%OS")) %>%
              ungroup() %>%
              mutate(f = row_number()) %>%
              dplyr::select(-filename)

            # assign id to each plant
            coord_df <- coordinates(plant_taxa_sp) %>%
              as_tibble() %>%
              mutate(id = row_number())

            # join data
            ps_df <- ps_mat %>%
              as_tibble() %>%
              left_join(time_df, by = "f") %>%
              left_join(coord_df, by = "id") %>%
              mutate(
                red = red * 0.0001, # scaling following Dixon et al's code
                green = green * 0.0001,
                blue = blue * 0.0001,
                nir = nir * 0.0001
              ) %>%
              # mutate(evi=2.5* (nir-red) / (nir + 6*red - 7.5*blue + 1),
              #        gndvi=(nir-green)/(nir+green),
              #        ebi= (red + green + blue) / (green / blue * (red - blue + 1))) %>%
              left_join(ps_mask_mat %>% as_tibble(), by = c("id", "f")) %>%
              dplyr::select(-f)

            # save
            write_rds(ps_df, paste0(ps_path, "analyses/ps_", siteoi, "_", taxaoi_short, ".rds"))
          }
        }
      }
    }
  )

  if (class(iserror) != "try-error") {
    iscomplete <- T
  } else if (class(iserror) == "try-error") { # restart cluster
    iscomplete <- F
    closeAllConnections()
    cl <- makeCluster(50, outfile = "")
    registerDoSNOW(cl)
  }
}
stopCluster(cl)
```

Example time series.
```{r, eval=TRUE}
site_vis <- "DT"
taxa_vis <- "Quercus"
set.seed(1)
read_rds(paste0(ps_path, "analyses/ps_", site_vis, "_", taxa_vis, ".rds")) %>%
  filter(id %in% ((.) %>% pull(id) %>% sample(4))) %>% # four random trees
  filter(qa == 0) %>%
  gather(key = "band", value = "value", -id, -time, -lon, -lat, -qa) %>%
  ggplot() +
  geom_point(aes(x = time, y = value, col = band), alpha = 0.25) +
  facet_wrap(. ~ id, ncol = 1) +
  theme_classic() +
  ggtitle(paste0("Taxa: ", taxa_vis, "; Site: ", site_vis))
```

## 1.6 HLS
<!-- Manually downloaded HLS_L30 from  -->
<!-- https://search.earthdata.nasa.gov/search/granules?p=C2021957657-LPCLOUD&pg[0][v]=f&pg[0][gsk]=-start_date&q=hls&tl=1652353102.502!3!! -->

<!-- bounding box from plant extent -->

<!-- Use api? -->

# 2 Correlation betwen leafing and flowering phenology on individual level
## 2.1 Preparation

```{r, eval=TRUE, echo=FALSE}
# Function for smoothing, because ptw::whit1 does not take NA in the time series
whitfun <- function(x, lambda) {
  max_id <- 0
  done <- F
  while (!done) {
    min_id <- min(which(!is.na(x[(max_id + 1):length(x)]))) + (max_id) # first number that is not NA
    if (min_id == Inf) { # all numbers are NA
      done <- T # consider this ts done
    } else {
      max_id <- min(which(is.na(x[min_id:length(x)]))) - 1 + (min_id - 1) # last number in the first consecutive non-NA segment
      if (max_id == Inf) {
        max_id <- length(x) # last non-NA segment is at the end of the whole ts
        done <- T # consider this ts done
      }
      x[min_id:max_id] <- ptw::whit1(x[min_id:max_id], lambda) # whitman smoothing for this non-NA segment
    }
  }
  return(x)
}
```


Some settings.
```{r, eval=TRUE}
# set color palette
cols <- c("EVI (PS)" = "dark green", "G2R (PS)" = "yellow green", "EBI (PS)" = "orange", "pollen count (NAB)" = "dark red", "flower observation (USA-NPN)" = "dark orchid", "flowering frequency" = "dark blue", "flower observation (DT)" = "coral")

# possible green up and green down thresholds
thres_list_up <- seq(from = 0, to = 1, by = 0.1) %>% round(1)
thres_list_down <- seq(from = 1, to = 0.0, by = -0.1) %>% round(1)
thres_df <- bind_rows(
  data.frame(direction = "up", threshold = thres_list_up),
  data.frame(direction = "down", threshold = thres_list_down)
)

# read in manually determined window of flowering
flower_window_df <- read_csv("/data/ZHULAB/phenology/nab/flower_window.csv")
```

Is there even a typical phenological curve? This is a function to rule out curves that are flat.
```{r, eval=TRUE}
flat_better <- function(ts, doy = (274 - 365):(365 + 151), k = 50) {
  fit_list <- vector(mode = "list")

  # fit simple linear regression model
  fit0 <- lm(ts ~ doy, data = data.frame(doy, ts))
  fit_list[[1]] <- data.frame(AIC = AIC(fit0, k = k), model = "fit0")

  # fit piecewise regression model to original model with different numbers of breakpoints
  try( # sometimes fitting may fail
    {
      fit1 <- segmented(fit0, seg.Z = ~doy, npsi = 1, it = 10, control = seg.control(seed = 42, fix.npsi = T))
      fit_list[[2]] <- data.frame(AIC = AIC(fit1, k = k), model = "fit1")
    },
    silent = T
  )

  try(
    {
      fit2 <- segmented(fit0, seg.Z = ~doy, npsi = 2, it = 10, control = seg.control(seed = 42, fix.npsi = T))
      fit_list[[3]] <- data.frame(AIC = AIC(fit2, k = k), model = "fit2")
    },
    silent = T
  )

  try(
    {
      fit3 <- segmented(fit0, seg.Z = ~doy, npsi = 3, it = 10, control = seg.control(seed = 42, fix.npsi = T))
      fit_list[[4]] <- data.frame(AIC = AIC(fit3, k = k), model = "fit3")
    },
    silent = T
  )

  # rank by AIC. k=50 sets a large penalty for more complex models
  aic_df <- bind_rows(fit_list) %>%
    arrange(AIC, model)

  # return 1 when straight line is the best fit
  better <- aic_df %>%
    head(1) %>%
    pull(model) == "fit0"
  return(better)
}
```

## 2.2 Compare with individual flowering observations (Detroit oaks) 
Trees in detroit.
```{r, eval=TRUE}
taxaoi <- "Quercus"
siteoi <- "DT"
yearoi <- 2017

taxaoi_short <- str_split(taxaoi, " ", simplify = T)[1]
flower_window <- seq(flower_window_df %>% filter(taxa == taxaoi) %>% pull(start),
  flower_window_df %>% filter(taxa == taxaoi) %>% pull(end),
  by = 1
)
thres_df_taxa <- thres_df %>% filter(direction == "up")
plant_taxa_df <- plant_df %>%
  filter(site == siteoi) %>%
  filter(genus == taxaoi_short | family == taxaoi_short) %>%
  mutate(id = row_number()) %>%
  drop_na(lon, lat)

# plant_df %>%
#   filter(site==siteoi) %>%
#   filter(genus==taxaoi_short|family==taxaoi_short) %>%
#   group_by(species) %>%
#   summarise(n=n()) %>%
#   ungroup() %>%
#   arrange(desc(n))

plant_taxa_df %>%
  ggplot() +
  geom_point(aes(x = lon, y = lat, col = species)) +
  theme_classic()
```
Flowering observations.
```{r, eval=TRUE}
detroit_df <- read_csv("/data/ZHULAB/phenology/occurrence/StreetTrees/Detroit_oak_pheno_obs_spring_2017.csv")

detroit_df_ts <- detroit_df %>%
  left_join(detroit_df %>%
    distinct(tree) %>%
    mutate(id = row_number()),
  by = "tree"
  ) %>%
  left_join(trees_df %>% filter(site == "DT") %>% dplyr::select(id, lon, lat), by = c("tree" = "id")) %>% # trees_df uses "tree" as "id"
  left_join(detroit_df %>%
    group_by(tree) %>%
    arrange(julian_day) %>%
    summarize(
      mindoy = min(julian_day),
      # peak=findpeaks(flowering_interp, sortstr = T)[1,2] %>% as.numeric(),
      thres = which(flowering_interp >= 95) %>% min() # first day that reaches 95% flowering 
    ) %>%
    mutate(peak = thres + mindoy - 1) %>% # correct with starting doy
    ungroup() %>%
    dplyr::select(tree, peak), by = c("tree")) %>%
  arrange(peak) %>%
  mutate(id = as.character(id)) %>%
  mutate(date = as.Date("2017-01-01") + julian_day - 1) %>%
  mutate(peak_date = as.Date("2017-01-01") + peak - 1)

detroit_df_ts %>%
  filter(id %in% ((.) %>% distinct(id) %>% pull(id) %>% sample(6))) %>%
  ggplot() +
  geom_point(aes(x = date, y = flowering_raw), col = "coral") +
  geom_line(aes(x = date, y = flowering_interp)) +
  geom_vline(aes(xintercept = peak_date), col = "dark orchid") +
  facet_wrap(. ~ id, ncol = 1) +
  ylab("flowering status")+
  theme_classic()
```

View EVI, NAB, and NPN data together for one tree.
```{r, eval=TRUE}
# preprocess ps data
ps_df <- read_rds(paste0(ps_path, "analyses/ps_", siteoi, "_", taxaoi_short, ".rds"))
ps_df_proc <- ps_df %>%
  drop_na() %>%
  mutate(date = as.Date(time)) %>%
  mutate(
    year = format(time, "%Y") %>% as.integer(),
    doy = format(time, "%j") %>% as.integer(),
    hour = format(strptime(time, "%Y-%m-%d %H:%M:%S"), "%H") %>% as.integer()
  ) %>%
  filter(qa == 0) %>%
  group_by(id, lon, lat, date, year, doy) %>%
  summarise(
    blue = mean(blue),
    green = mean(green),
    red = mean(red),
    nir = mean(nir)
  ) %>%
  ungroup() %>%
  mutate(evi = 2.5 * (nir - red) / (nir + 6 * red - 7.5 * blue + 1)) %>%
  filter(evi > 0, evi <= 1) %>%
  filter(red > 0, green > 0, blue > 0)

# subset nab data
pollen_df <- nab_with_taxa_df %>%
  left_join(meta_df %>% dplyr::select(id, site), by = "id") %>%
  filter(site == siteoi) %>%
  filter(genus == taxaoi_short | family == taxaoi_short)

# subset npn data
npn_df <- npn_df_all %>%
  filter(site == siteoi) %>%
  filter(taxa == taxaoi_short)

# join ps, nab, and npn data
ts_df <- ps_df_proc %>%
  left_join(plant_taxa_df, by = c("id", "lon", "lat")) %>%
  dplyr::select(id, date,
    `EVI (PS)` = evi 
  ) %>%
  mutate(id = as.factor(id)) %>%
  full_join(pollen_df %>%
    dplyr::select(date, `pollen count (NAB)` = count) %>%
    mutate(id = "pollen"),
  by = c("date", "id")
  ) %>%
  full_join(npn_df %>%
    dplyr::select(date, `flower observation (USA-NPN)` = count) %>%
    mutate(id = "npn"),
  by = c("date", "id")
  ) %>%
  arrange(id, date) %>%
  mutate(doy = format(date, "%j") %>% as.numeric()) %>%
  mutate(year = format(date, "%Y") %>% as.numeric())

# focus on one year, spanning from day 274 in the previous year to day 151 in the next year
ts_df_subset <- ts_df %>%
  filter(doy != 366) %>%
  # filter(doy>start_doy,doy<=end_doy) %>%
  filter(year == yearoi | year == (yearoi - 1) | year == (yearoi + 1)) %>%
  mutate(doy = ifelse(doy > 273 & year == yearoi - 1, doy - 365, doy)) %>%
  mutate(year = ifelse(doy <= 0 & year == yearoi - 1, year + 1, year)) %>%
  mutate(doy = ifelse(doy < 152 & year == yearoi + 1, doy + 365, doy)) %>%
  mutate(year = ifelse(doy > 365 & year == yearoi + 1, year - 1, year)) %>%
  filter(year == yearoi) %>%
  gather(key = "var", value = "value", -date, -id, -doy, -year) %>%
  mutate(var = fct_relevel(var, levels = c("EVI (PS)", "G2R (PS)", "EBI (PS)", "pollen count (NAB)", "flower observation (USA-NPN)"))) %>% 
  arrange(doy)

# join with flowering data
ts_df_subset <- ts_df_subset %>%
  bind_rows(detroit_df_ts %>% dplyr::select(id, date, lon, lat, value = flowering_raw, doy = julian_day) %>%
    mutate(
      var = "flower observation (DT)",
      year = 2017
    ))

# summarize quantiles on the site level
ts_df_subset_summary <- ts_df_subset %>%
  drop_na(value) %>%
  group_by(date, var, doy, year) %>%
  summarise(
    q1 = quantile(value, 0.05, na.rm = T),
    q2 = quantile(value, 0.5, na.rm = T),
    q3 = quantile(value, 0.95, na.rm = T),
    n = n()
  ) %>%
  filter(n > 1) %>%
  ungroup()

# visualize for one tree
idoi="1"
plant_sp<-plant_taxa_df %>% filter(id==idoi) %>% pull(species)
ts_df_subset %>%
  filter(id == idoi | id == "pollen" | id == "npn") %>%
  ggplot() +
  geom_point(aes(x = date, y = value, col = var)) +
  facet_wrap(. ~ var, ncol = 1, scales = "free_y") +
  scale_color_manual(values = cols) +
  ylab("")+
  ggtitle(paste0("Site: ", siteoi,"  Year: ", yearoi,"  ID: ", idoi, "  Species: ", plant_sp))+
  theme_classic()
```

Detect greenup at multiple thresholds.
```{r, eval=TRUE}
get_doy<-function (plant_df, thres_df, ts_df, idoi) {
  plant_sp<-plant_df %>% filter(id==idoi) %>% pull(species)
  px_doy<-ts_df %>% filter(id==idoi) %>% filter(var=="EVI (PS)") %>% pull(doy)
  
  if (length(px_doy [px_doy>=1 & px_doy<=365])<50) {
    flower_df_up<-flower_df_down<-NULL
    print("too few data points")
  }  else {
    px_evi<-ts_df %>% filter(id==idoi) %>% filter(var=="EVI (PS)") %>% pull(value)
    px_evi_in<-approx(px_doy, px_evi, (274-365):(365+151), rule=1)$y
    
    px_evi_in_sm<-whitfun(px_evi_in,30)
    
    flatbetter<-flat_better(px_evi_in_sm, k=50)
    
    ### green down
    thres_list_down<-thres_df %>% filter(direction=="down") %>% pull(threshold)
    if (length(thres_list_down)==0) {
      flower_df_down<-NULL
    } else {
      max_evi<-quantile(px_evi_in_sm[(-274+365+1):(-274+365+365+1)], 1, na.rm = T)
      start_doy<-which(!is.na(px_evi_in_sm[(-274+365+1):(-274+365+365+1)])&px_evi_in_sm[(-274+365+1):(-274+365+365+1)]>=max_evi) %>% max()-274+365
      min_evi<-quantile(px_evi_in_sm[start_doy:length(px_evi_in_sm)], 0, na.rm = T)
      min_doy<-which(!is.na(px_evi_in_sm[start_doy:length(px_evi_in_sm)])& px_evi_in_sm[start_doy:length(px_evi_in_sm)]<=min_evi) %>% min()+start_doy
      end_doy<-min_doy
      
      param_ok2<- (end_doy>start_doy) & (!flatbetter)
      
      if (!param_ok2) {
        greendown_doy<-rep(NA, length(thres_list_down))
        start_doy<-NA
        end_doy<-NA
        print("not typical growth curve")
      } else {
        greendown_thres<-(max_evi-min_evi)*thres_list_down+min_evi
        greendown_doy<-rep(NA, length(greendown_thres))
        for (t in 1:length(greendown_thres)) {
          greendown_doy[t]<-which(px_evi_in_sm[start_doy:end_doy]<=greendown_thres[t]) %>% min()+start_doy
        }
        start_doy=start_doy+274-365
        end_doy=end_doy+274-365
        greendown_doy=greendown_doy+274-365
      }
      flower_df_down<-data.frame(id=idoi, start=start_doy, end=end_doy,direction="down", thres=thres_list_down, doy=greendown_doy) 
    }
    
    ### green up
    thres_list_up<-thres_df %>% filter(direction=="up") %>% pull(threshold)
    if (length(thres_list_up)==0) {
      flower_df_up<-NULL
    } else {
      max_evi<-quantile(px_evi_in_sm[(-274+365+1):(-274+365+365+1)], 1, na.rm = T)
      end_doy<-which(!is.na(px_evi_in_sm[(-274+365+1):(-274+365+365+1)])&px_evi_in_sm[(-274+365+1):(-274+365+365+1)]>=max_evi) %>% min()-274+365
      min_evi<-quantile(px_evi_in_sm[1:end_doy], 0.00, na.rm = T)
      min_doy<-which(!is.na(px_evi_in_sm[1:end_doy])& px_evi_in_sm[1:end_doy]<=min_evi) %>% max()
      start_doy<-min_doy
      
      param_ok2<- (end_doy>start_doy) & (!flatbetter)
      
      if (!param_ok2) {
        greenup_doy<-rep(NA, length(thres_list_up))
        start_doy<-NA
        end_doy<-NA
        print("not typical growth curve")
      } else {
        greenup_thres<-(max_evi-min_evi)*thres_list_up+min_evi
        greenup_doy<-rep(NA, length(greenup_thres))
        for (t in 1:length(greenup_thres)) {
          greenup_doy[t]<-which(px_evi_in_sm[start_doy:end_doy]>=greenup_thres[t]) %>% min()+start_doy 
        }
        start_doy=start_doy+274-365
        end_doy=end_doy+274-365
        greenup_doy=greenup_doy+274-365
      }
      flower_df_up<-data.frame(id=idoi, start=start_doy, end=end_doy,direction="up", thres=thres_list_up, doy=greenup_doy) 
    }
  }
  
  flower_df<-bind_rows(flower_df_up, flower_df_down)
  
  return (flower_df)
}

```

```{r, eval=TRUE}
plot_tree<-function (plant_df, ts_df, flower_df, idoi)  {
  plant_sp<-plant_df %>% filter(id==idoi) %>% pull(species)
  
    p_1tree<-ggplot( )+
      geom_point(data=ts_df_subset %>% filter(id==idoi|id=="npn"|id=='pollen') %>% 
                   filter(var %in% c("EVI (PS)","pollen count (NAB)", "flower observation (USA-NPN)", "flower observation (DT)")),
                 aes(doy, value,group=as.factor(id), col=var))+
      theme_classic()+
      guides(col=F,
             alpha=F)+
      scale_color_manual(values=cols)+
      facet_wrap(.~var, ncol=1, scales = "free_y")+
      xlab("day of year")+
      ylab("")+
      ggtitle(paste0("Site: ", siteoi,"  Year: ", yearoi,"  ID: ", idoi, "  Species: ", plant_sp))
    if (nrow(flower_df)==0) {
      p_1tree
    } else {
      p_1tree<-p_1tree+
        geom_vline(data=flower_df, aes(xintercept = start), col="red", alpha=0.8)+
        geom_vline(data=flower_df, aes(xintercept = end), col="red", alpha=0.8)+
        geom_vline(data=flower_df, aes(xintercept = doy), col="red", alpha=0.3)
    }
    
  return (p_1tree)
}
```

```{r, eval=TRUE}
idoi <- "1"
flower_df <- get_doy(plant_taxa_df, thres_df_taxa,ts_df_subset, idoi)
p_1tree <- plot_tree(plant_taxa_df, ts_df_subset, flower_df, idoi)
p_1tree
```

Repeat for all trees to get frequency distribution of green-up dates.
```{r, eval=TRUE}
# get green-up/green-down doy for each tree
random_id <- unique(plant_taxa_df$id)
flower_df_list <- vector(mode = "list", length = length(random_id))
for (i in 1:length(random_id)) {
  blas_set_num_threads(1)
  omp_set_num_threads(1)

  idoi <- as.character(random_id)[i]

  flower_df <- get_doy(plant_taxa_df, thres_df_taxa,ts_df_subset, idoi)

  # print(paste0(i , " out of ", length(random_id)))
  flower_df_list[[i]] <- flower_df
}

# bind with species and site info
flower_doy_df <- bind_rows(flower_df_list) %>%
  left_join(plant_taxa_df %>% dplyr::select(id, species, lat, lon) %>% mutate(id = as.character(id)), by = "id") %>%
  mutate(site = siteoi, year = yearoi)

# summarize frequency on the site level
flower_freq_df_list <- vector(mode = "list", length = nrow(thres_df_taxa))
for (t in 1:nrow(thres_df_taxa)) {
  flower_freq_df_list[[t]] <- flower_doy_df %>%
    drop_na(start, end) %>%
    filter(
      direction == thres_df_taxa$direction[t],
      thres == thres_df_taxa$threshold[t]
    ) %>%
    group_by(doy, thres, direction) %>%
    summarise(count = n()) %>%
    ungroup() %>%
    mutate(freq = count / n())
}
flower_freq_df <- bind_rows(flower_freq_df_list)

# join frequency distribution with EVI, NAB and NPN data
flower_freq_df <- flower_freq_df %>%
  mutate(doy = factor(doy, levels = c((274 - 365):(365 + 151)))) %>%
  complete(doy, thres, direction, fill = list(count = 0, freq = 0)) %>%
  mutate(doy = doy %>% as.character() %>% as.numeric()) %>%
  arrange(doy) %>%
  full_join(ts_df_subset %>%
    filter(id == "pollen") %>%
    filter(var == "pollen count (NAB)") %>%
    filter(year == yearoi) %>%
    dplyr::select(doy, pollen = value),
  by = "doy"
  ) %>%
  full_join(ts_df_subset %>%
    filter(id == "npn") %>%
    filter(var == "flower observation (USA-NPN)") %>%
    filter(year == yearoi) %>%
    dplyr::select(doy, npn = value),
  by = "doy"
  ) %>%
  full_join(ts_df_subset_summary %>%
    filter(var == "EVI (PS)") %>%
    filter(year == yearoi) %>%
    dplyr::select(doy, evi = q2),
  by = "doy"
  ) %>%
  mutate(doy = as.numeric(doy)) %>%
  mutate(site = siteoi, year = yearoi) %>%
  drop_na(doy, thres)
```

Plot green-up doy at several thresholds vs. flower observations. (Only showing within the limit of day 110 to 160.)
```{r}
flower_doy_df %>%
  filter(thres %in% c(0.4, 0.5, 0.6, 0.7, 0.8)) %>%
  left_join(detroit_df_ts %>% dplyr::distinct(id, peak) %>% filter(is.finite(peak)), by = "id") %>%
  ggplot() +
  geom_jitter(aes(x = peak, y = doy, group = interaction(direction, thres), col = interaction(direction, thres)), alpha = 1) +
  geom_smooth(aes(x = peak, y = doy, group = interaction(direction, thres), col = interaction(direction, thres)), method = "lm") +
  geom_abline(slope = 1, intercept = 0, col = "red") +
  ggpubr::stat_cor(
    aes(x = peak, y = doy, group = interaction(direction, thres),
        col = interaction(direction, thres),
        label = paste(..r.label.., ..p.label.., sep = "*`,`~")),
    p.accuracy = 0.05,
        label.x.npc = 0.7,
        label.y.npc = "top",
        show.legend=F
  )+
  coord_equal() +
  ylim(c(110, 160)) +
  xlim(c(110, 160)) +
  xlab ("flowering doy")+
  ylab ("green-up doy")+
  labs(col="direction and threshold")+
  theme_classic()
```

Repeat for all years and check correlation. Here only showing green-up at 50%. There were significantly positive correlations in three out of five years, suggesting that green-up and flowering time may be correlated on the individual level.
```{r, eval=TRUE}
read_rds("/raid/users/ysong67/GitHub/phenology/RS4flower/output/data/Quercus/flowering day of year.rds") %>% 
  filter(site==siteoi) %>%
  bind_rows(flower_doy_df %>% mutate(year=yearoi)) %>% 
  filter(thres==0.5,
         direction=="up") %>% 
  left_join(detroit_df_ts %>% dplyr::distinct(id, peak)  %>% filter(is.finite(peak)), by="id") %>% 
  ggplot()+
  geom_jitter(aes(x=peak, y=doy, group=as.factor(year), col=as.factor(year)), alpha=1)+
  geom_smooth(aes(x=peak, y=doy, group=as.factor(year), col=as.factor(year)), method="lm")+
  geom_abline(slope = 1, intercept = 0, col="red")+
  ggpubr::stat_cor(
    aes(x = peak, y = doy, group = as.factor(year),
        col = as.factor(year),
        label = paste(..r.label.., ..p.label.., sep = "*`,`~")),
    p.accuracy = 0.05,
        label.x.npc = 0.7,
        label.y.npc = "top",
        show.legend=F
  )+
  coord_equal()+
  ylim(c(110,160 ))+
  xlim(c(110,160 ))+
  theme_classic()
```

Compare maps.
```{r}
read_rds("/raid/users/ysong67/GitHub/phenology/RS4flower/output/data/Quercus/flowering day of year.rds") %>% 
  filter(site==siteoi) %>%
  bind_rows(flower_doy_df %>% mutate(year=yearoi)) %>% 
  filter(thres==0.5,
         direction=="up") %>% 
  dplyr::select(id, doy, lat, lon,year) %>% 
  mutate(year=as.factor(year)) %>% 
  bind_rows(detroit_df_ts %>% rename(doy=peak) %>% dplyr::distinct(id, doy, lat, lon) %>% filter(is.finite(doy)) %>% mutate(year="flower observation (DT)")) %>% 
    filter(doy<=160 & doy>=110) %>% 
  spread(key="year", value="doy") %>% 
  # drop_na() %>% 
  rowwise() %>% 
  mutate(lat=lat+rnorm(1, 0, 0.002),
         lon=lon+rnorm(1, 0, 0.002)) %>% 
  ungroup() %>% 
  gather(key="group", value="doy", -id, -lat, -lon)  %>% 
  group_by(group) %>% 
  mutate(doy=(doy-mean(doy, na.rm = T))/(quantile(doy, 0.95, na.rm = T)-quantile(doy, 0.05, na.rm = T))) %>% # standardize to between -0.5 and 0.5
  filter(doy>=-0.5 & doy<=0.5) %>% # remove outliers
  ungroup() %>% 
  ggplot()+
  geom_point(aes(x=lon, y=lat, col=doy))+
  facet_wrap(.~group, ncol=2)+
  coord_equal()+
  theme_classic()+
  scale_color_viridis_c(direction = -1)
```

apply tp hls in dt

# 3 Match pollen phenology with leafing phenology
Conceptual figure.
```{r, eval=TRUE}
conceptual1<-image_read("/raid/users/ysong67/GitHub/phenology/RS4flower/output/figures/conceptual1.jpg")
p_conceptual1<-image_ggplot(conceptual1)
conceptual2<-image_read("/raid/users/ysong67/GitHub/phenology/RS4flower/output/figures/conceptual1.jpg")
p_conceptual2<-image_ggplot(conceptual2)
grid.arrange(p_conceptual1, p_conceptual2, ncol=1)
```

## 3.1 Characterize leafing phenology for all taxa and cities
```{r, eval=FALSE}
cl <- makeCluster(50, outfile = "")
registerDoSNOW(cl)

for (taxaoi in taxa_list) {
  taxaoi_short<-str_split(taxaoi, " ", simplify = T)[1]
  flower_window<-seq(flower_window_df %>% filter(taxa==taxaoi) %>% pull(start),
                     flower_window_df %>% filter(taxa==taxaoi) %>% pull(end),
                     by=1)
  if (taxaoi %in% c("Ambrosia", "Ulmus late")) {
    thres_df_taxa<-thres_df %>% filter(direction=="down")
  } else if (taxaoi== "Poaceae early"  ) {
    thres_df_taxa<-thres_df %>% filter(threshold>=0.5|direction=="up")
  } else if (taxaoi== "Poaceae late"  ) {
    thres_df_taxa<-thres_df %>% filter(threshold>=0.5|direction=="down")
  } else {
    thres_df_taxa<-thres_df %>% filter(direction=="up")
  }
  flower_doy_df_siteyears_list<-flower_freq_df_siteyears_list<-vector(mode="list", length=length(site_list))
  for (s in 1:length(site_list)) {
    siteoi<-site_list[s]
    plant_taxa_df<-plant_df %>% 
      filter(site==siteoi) %>% 
      filter(genus==taxaoi_short|family==taxaoi_short) %>% 
      mutate(id=row_number()) %>% 
      drop_na(lon, lat)
    
    plant_df %>%
      filter(site==siteoi) %>%
      filter(genus==taxaoi_short|family==taxaoi_short) %>%
      group_by(species) %>%
      summarise(n=n()) %>%
      ungroup() %>%
      arrange(desc(n))
    
    if (taxaoi %in% c("Ambrosia","Poaceae")) {
      min_sample_size=10
    } else {
      min_sample_size=20
    }
    
    if (nrow(plant_taxa_df)>=min_sample_size) {
      set.seed(1)
      random_id<<-sample(plant_taxa_df$id, min(2000, length(unique(plant_taxa_df$id)))) %>% sort()
      
      # preprocess ps data
      ps_df<-read_rds(paste0(ps_path, "analyses/ps_",siteoi,"_",taxaoi_short,".rds")) 
      ps_df_proc<- ps_df%>% 
        drop_na() %>% 
        filter(id%in% random_id) %>%
        mutate(date=as.Date(time)) %>% 
        mutate(year=format(time, "%Y") %>% as.integer(),
               doy=format(time, "%j") %>% as.integer(),
               hour=format(strptime(time,"%Y-%m-%d %H:%M:%S"),'%H') %>% as.integer()) %>% 
        filter(qa==0) %>%
        group_by(id, lon, lat, date, year, doy ) %>% 
        summarise(blue=mean(blue),
                  green=mean(green),
                  red=mean(red),
                  nir=mean(nir)) %>% 
        ungroup() %>% 
        mutate(evi=2.5* (nir-red) / (nir + 6*red - 7.5*blue + 1)) %>% 
        filter(evi>0, evi<=1) %>% 
        filter(red>0, green>0, blue>0) 
      
      # subset nab data
      pollen_df<- nab_with_taxa_df %>% 
        left_join(meta_df %>% dplyr::select(id, site), by="id") %>% 
        filter(site==siteoi) %>% 
        filter(genus==taxaoi_short|family==taxaoi_short) 
      
      # subset npn data
      npn_df<-npn_df_all %>% 
        filter(site==siteoi) %>% 
        filter(taxa==taxaoi_short)
      
      ts_df<-ps_df_proc %>% 
        left_join(plant_taxa_df, by=c("id", "lon", "lat")) %>% 
        
        dplyr::select(id, date, `EVI (PS)`=evi #, 
                      # `G2R (PS)`=g2r, `EBI (PS)`=ebi
                      ) %>% 
        mutate(id=as.factor(id)) %>% 
        full_join(pollen_df %>% 
                    dplyr::select(date, `pollen count (NAB)`=count) %>% 
                    mutate(id="pollen") , 
                  by=c("date", "id") ) %>% 
        full_join(npn_df %>%
                    dplyr::select(date, `flower observation (USA-NPN)`=count) %>%
                    mutate(id="npn") ,
                  by=c("date", "id") ) %>%
        arrange(id, date) %>% 
        mutate(doy=format(date, "%j") %>% as.numeric()) %>% 
        mutate(year=format(date, "%Y") %>% as.numeric()) 
      
      flower_doy_df_years_list<-flower_freq_df_years_list<-vector(mode="list", length=length(year_list))
      for (y in 1:length(year_list)) {
        yearoi<-year_list[y]
        ts_df_subset<-ts_df %>% 
          filter(doy!=366) %>% 
          filter(year==yearoi|year==(yearoi-1)|year==(yearoi+1)) %>% 
          mutate(doy=ifelse(doy>273& year==yearoi-1, doy-365, doy)) %>% 
          mutate(year=ifelse(doy<=0 & year==yearoi-1, year+1, year)) %>% 
          mutate(doy=ifelse(doy<152& year==yearoi+1, doy+365, doy)) %>% 
          mutate(year=ifelse(doy>365 & year==yearoi+1, year-1, year)) %>% 
          filter(year==yearoi) %>% 
          gather(key="var", value="value", -date, -id, -doy, -year ) %>% 
          mutate(var=fct_relevel(var, levels=c( "EVI (PS)", "G2R (PS)","EBI (PS)", "pollen count (NAB)", "flower observation (USA-NPN)"))) %>% 
          mutate(alpha=case_when(var %in% c("EVI", "G2R")~0.05,
                                 TRUE~1)) %>% 
          arrange(doy)
        
        ts_df_subset_summary<- ts_df_subset%>%
          drop_na(value) %>%
          group_by(date, var, doy, year) %>%
          summarise(q1=quantile(value, 0.05, na.rm=T),
                    q2=quantile(value, 0.5, na.rm=T),
                    q3=quantile(value, 0.95, na.rm=T),
                    n=n()) %>%
          filter(n>1) %>%
          ungroup()
        
        flower_df_list<-
          foreach (i = 1:length(random_id),
                   .packages = c("tidyverse", "ptw","greenbrown", "EnvCpt","segmented", "RhpcBLASctl"))  %dopar% {
                     blas_set_num_threads(1)
                     omp_set_num_threads(1)
                     
                     debug=F
                     idoi <-  as.character(random_id)[i]
                     
                     if (debug) { 
                       set.seed(NULL)
                       idoi<-sample(random_id,1) %>% as.character()
                     }
                     
                     flower_df<-get_doy(plant_taxa_df,thres_df_taxa, ts_df_subset, idoi)
                     
                     if (debug) {
                       p_1tree<-plot_tree(plant_taxa_df, ts_df_subset, flower_df, idoi)
                     }
                     
                     print(paste0(i , " out of ", length(random_id)))
                     flower_df
                   }
        flower_doy_df<-bind_rows(flower_df_list) %>% 
          left_join(plant_taxa_df %>% dplyr::select(id, species, lat, lon) %>% mutate(id=as.character(id)), by="id") %>% 
          mutate(site=siteoi, year=yearoi)
        
        flower_freq_df_list<-vector(mode="list", length=nrow(thres_df_taxa))
        for (t in 1:nrow(thres_df_taxa)) {
          flower_freq_df_list[[t]]<-flower_doy_df %>% 
            drop_na(start, end) %>% 
            filter(direction==thres_df_taxa$direction[t],
                   thres==thres_df_taxa$threshold[t]) %>% 
            group_by(doy, thres,direction) %>%
            summarise(count=n()) %>%
            ungroup() %>% 
            mutate(freq=count/n())
        }
        flower_freq_df<-bind_rows(flower_freq_df_list)
        
        
        flower_freq_df<-flower_freq_df %>% 
          mutate(doy=factor(doy, levels=c((274-365):(365+151))) ) %>% 
          complete(doy, thres,direction,fill=list(count=0,freq=0)) %>%
          mutate(doy=doy %>% as.character() %>% as.numeric()) %>% 
          arrange(doy) %>% 
          full_join(ts_df_subset %>% 
                      filter(id=="pollen") %>% 
                      filter(var=="pollen count (NAB)") %>% 
                      filter(year==yearoi) %>% 
                      dplyr::select( doy, pollen=value),
                    by="doy") %>% 
          full_join(ts_df_subset %>%
                      filter(id=="npn") %>%
                      filter(var=="flower observation (USA-NPN)") %>%
                      filter(year==yearoi) %>%
                      dplyr::select( doy, npn=value),
                    by="doy") %>%
          full_join(ts_df_subset_summary %>% 
                      filter(var=="EVI (PS)") %>% 
                      filter(year==yearoi) %>% 
                      dplyr::select( doy, evi=q2),
                    by="doy"
          ) %>% 
          full_join(ts_df_subset_summary %>% 
                      filter(var=="G2R (PS)") %>% 
                      filter(year==yearoi) %>% 
                      dplyr::select( doy, g2r=q2),
                    by="doy"
          ) %>% 
          mutate(doy=as.numeric(doy)) %>% 
          mutate(site=siteoi, year=yearoi) %>% 
          drop_na(doy, thres)
        
        print(paste0(siteoi, ", ", yearoi))
        
        flower_doy_df_years_list[[y]]<-flower_doy_df 
        flower_freq_df_years_list[[y]]<-flower_freq_df
      }
      flower_doy_df_siteyears_list[[s]]<-bind_rows(flower_doy_df_years_list)
      flower_freq_df_siteyears_list[[s]]<-bind_rows(flower_freq_df_years_list)
    }
  }
  flower_doy_df<-bind_rows(flower_doy_df_siteyears_list) %>% 
    left_join(data.frame(site=site_list, sitename=sitename_list
                         # , region=region_list
                         ), by=c("site"))
  
  flower_freq_df<-bind_rows(flower_freq_df_siteyears_list) %>% 
    group_by(site, year,thres,direction) %>% 
    mutate(freq_sm=whit1(freq, 30)) %>% 
    mutate(pollen_in=na.approx(pollen, doy, na.rm=F, maxgap=14)) %>%  
    mutate(pollen_fill=replace_na(pollen_in, 0)) %>%
    mutate(pollen_sm=whitfun(pollen_fill,30)) %>%  
    ungroup() %>% 
    dplyr::select(-pollen_in, -pollen_fill) %>% 
    mutate(pollen=case_when(doy %in% flower_window~ pollen)) %>% 
    mutate(pollen_sm=case_when(doy %in% flower_window~ pollen_sm,
                               TRUE~0)) %>% 
    mutate(npn=case_when(doy %in% flower_window~ npn))
  
  # pollen climatology (historical mean of smoothed pollen count)
  pollen_clim<-flower_freq_df %>% 
    group_by(site,direction, thres,doy) %>%
    summarise(pollen_clim=mean(pollen_sm, na.rm=T)) 
  
  flower_freq_df<-flower_freq_df%>% 
    left_join(pollen_clim, by=c("site",  "direction","thres", "doy")) %>% 
    left_join(data.frame(site=site_list, sitename=sitename_list
                         # , region=region_list
                         ), by=c("site"))
  
  
  output_path<-paste0("/raid/users/ysong67/GitHub/phenology/RS4flower/output/data/", taxaoi,"/")
  dir.create(output_path, recursive = T)
  write_rds(flower_doy_df,paste0(output_path,"flowering day of year.rds" ))
  write_rds(flower_freq_df,paste0(output_path,"flowering frequency.rds" ) )
}
stopCluster(cl)
```

## 3.2 Tune threshold and time lag with pollen data
```{r, eval=FALSE}
cl <- makeCluster(50, outfile = "")
registerDoSNOW(cl)
for (taxaoi in taxa_list) {
  taxaoi_short<-str_split(taxaoi, " ", simplify = T)[1]
  flower_window<-seq(flower_window_df %>% filter(taxa==taxaoi) %>% pull(start),
                     flower_window_df %>% filter(taxa==taxaoi) %>% pull(end),
                     by=1)
  if (taxaoi %in% c("Ambrosia", "Ulmus late")) {
    thres_df_taxa<-thres_df %>% filter(direction=="down")
  } else if (taxaoi== "Poaceae early"  ) {
    thres_df_taxa<-thres_df %>% filter(threshold>=0.5|direction=="up")
  } else if (taxaoi== "Poaceae late"  ) {
    thres_df_taxa<-thres_df %>% filter(threshold>=0.5|direction=="down")
  } else {
    thres_df_taxa<-thres_df %>% filter(direction=="up")
  }
  
  output_path<-paste0("/raid/users/ysong67/GitHub/phenology/RS4flower/output/data/", taxaoi,"/")
  
  # read in leafing phenology data
  flower_doy_df<-read_rds(paste0(output_path,"flowering day of year.rds" ))
  flower_freq_df<-read_rds(paste0(output_path,"flowering frequency.rds" ) )
  
  # standardize all data to be between 0 and 1
  flower_freq_df_standard<-flower_freq_df %>% 
    mutate(pollen=pollen %>% sqrt()) %>%
    mutate(pollen_sm=pollen_sm %>% sqrt()) %>%
    mutate(pollen_clim=pollen_clim %>% sqrt()) %>%
    group_by(site, sitename, thres) %>%
    mutate(freq=(freq-min(freq, na.rm = T))/(max(freq, na.rm = T)-min(freq, na.rm = T))) %>% 
    mutate(freq_sm=(freq_sm-min(freq_sm, na.rm = T))/(max(freq_sm, na.rm = T)-min(freq_sm, na.rm = T))) %>% 
    mutate(pollen=(pollen-min(pollen, na.rm = T))/(max(pollen, na.rm = T)-min(pollen, na.rm = T))) %>%
    mutate(pollen_clim=(pollen_clim-min(pollen_sm, na.rm = T))/(max(pollen_sm, na.rm = T)-min(pollen_sm, na.rm = T))) %>% 
    mutate(pollen_sm=(pollen_sm-min(pollen_sm, na.rm = T))/(max(pollen_sm, na.rm = T)-min(pollen_sm, na.rm = T))) %>% 
    mutate(npn=(npn-min(npn, na.rm = T))/(max(npn, na.rm = T)-min(npn, na.rm = T))) %>%
    mutate(evi=(evi-min(evi, na.rm = T))/(max(evi, na.rm = T)-min(evi, na.rm = T))) %>% 
    mutate(g2r=(g2r-min(g2r, na.rm = T))/(max(g2r, na.rm = T)-min(g2r, na.rm = T))) %>% 
    ungroup()
  
  # optimize threshold and lag
  flower_freq_df_check_list<-vector(mode="list", length=length(site_list))
  for (r in 1:length(site_list)) { # select a region, here each site is treated as an individual region
    regionoi<-site_list[r]
    flower_freq_df_check_region_list<-vector(mode="list", length=nrow(thres_df_taxa))
    
    for (t in 1:nrow(thres_df_taxa)) { # select a threshold
      flower_freq_df_standard_subset<-flower_freq_df_standard %>% 
        filter(site==regionoi) %>%
        filter(direction==thres_df_taxa$direction[t],
               thres==thres_df_taxa$threshold[t]) %>% 
        group_by(site, year) %>% 
        ungroup()
      sample_size<-flower_freq_df_standard_subset %>% filter(!is.na(pollen), pollen>0) %>% nrow()
      if (sample_size>=5*4) { # only do tuning when there are more than 20 none-zero pollen count. skip the taxa and site otherwise.
        pollen_sm_ts<- flower_freq_df_standard_subset %>% pull (pollen_sm) # smoothed pollen count, for tuning
        pollen_ts<- flower_freq_df_standard_subset %>% pull (pollen) # pollen count, for calculating accuracy
        
        pollen_clim_ts<- flower_freq_df_standard_subset %>% pull (pollen_clim) # climatology
        rmse_clim<-sqrt(mean((pollen_clim_ts-pollen_ts)^2, na.rm=T)) # rmse between climatology and pollen count
        
        lags_list<- -200:200 # possible lags to try
        flower_freq_df_check_thres_list<-
          foreach (l = 1:length(lags_list),
                   .packages = c("tidyverse")) %dopar% {
                     lag=lags_list[l]
                     if (lag<0) {
                       flower_freq_df_standard_subset_lag<- flower_freq_df_standard_subset %>% mutate(freq_sm=lead(freq_sm,n=-lag))  %>% mutate(freq_sm=replace_na(freq_sm, 0)) # shift leafing phenology earlier by "lag" days, fill the NA with 0
                     } else if (lag==0) {
                       flower_freq_df_standard_subset_lag<-flower_freq_df_standard_subset  %>% mutate(freq_sm=replace_na(freq_sm, 0))
                     } else if (lag>0) {
                       flower_freq_df_standard_subset_lag<-flower_freq_df_standard_subset %>% mutate(freq_sm=lag(freq_sm,n=lag)) %>% mutate(freq_sm=replace_na(freq_sm, 0)) # shift leafing phenology later by "lag" days, fill the NA with 0
                     }
                     freq_ts_lag<- flower_freq_df_standard_subset_lag %>% pull (freq_sm)
                     rmse<-sqrt(mean((freq_ts_lag-pollen_sm_ts)^2, na.rm=T)) # rmse between remotely-sensed flowering phenology and smoothed pollen count
                     rmse_ps<-sqrt(mean((freq_ts_lag-pollen_ts)^2, na.rm=T)) # rmse between remotely-sensed flowering phenology and pollen count
                     
                     print(paste0("region ", r, ", threshold ", t, ", lag ", l))
                     
                     data.frame(direction=thres_df_taxa$direction[t],thres=thres_df_taxa$threshold[t],lag=lag, rmse=rmse,rmse_ps=rmse_ps, rmse_clim=rmse_clim)
                   }
        flower_freq_df_check_region_list[[t]]<-bind_rows(flower_freq_df_check_thres_list) %>% 
          arrange(rmse) %>% # choose the lag giving the smallest rmse in the threshold
          head(1)
      } else {
        flower_freq_df_check_region_list[[t]]<-data.frame(thres=numeric(0),lag=numeric(0), rmse=numeric(0), rmse_ps=numeric(0),rmse_clim=numeric(0))
      }
    }
    flower_freq_df_check_list[[r]]<-bind_rows(flower_freq_df_check_region_list) %>% 
      mutate(region=regionoi)  %>%
      arrange((rmse))
  }
  flower_freq_df_check<-bind_rows(flower_freq_df_check_list)
  # flower_freq_df_check
  write_rds(flower_freq_df_check,paste0(output_path,"accuracy check.rds") )
  
  best_thres<-flower_freq_df_check %>% 
    group_by(direction, thres) %>% 
    summarise(rmse=median(rmse)) %>% # median rmse for each threshold
    arrange(rmse) %>% 
    head(1) %>% # keep threshold giving the smallest median rmse
    dplyr::select(direction,thres)
  
  # flower_freq_df_check %>% filter(direction==best_thres$direction[1],
  #                                 thres==best_thres$thres[1])
  
  # getting time series with best threshold and lag
  flower_freq_df_standard_best_list<-vector(mode="list", length=length(site_list))
  for (r in 1:length(site_list)) {
    regionoi<-site_list[r]
    lagoi<-flower_freq_df_check %>% filter(direction==best_thres$direction,thres==best_thres$thres, region==regionoi) %>% pull(lag)
    if (length(lagoi)>0) {
      flower_freq_df_standard_best<-flower_freq_df_standard %>% 
        filter(site==regionoi) %>% 
        filter(direction==best_thres$direction,
               thres == best_thres$thres) %>%
        ungroup() %>%
        group_by(site, year)
      
      if (lagoi<0) {
        flower_freq_df_standard_best<- flower_freq_df_standard_best %>% mutate(freq_sm=lead(freq_sm,n=-lagoi)) %>% mutate(freq_sm=replace_na(freq_sm, 0))
      } else if (lagoi==0) {
        flower_freq_df_standard_best<- flower_freq_df_standard_best %>% mutate(freq_sm=replace_na(freq_sm, 0))
      } else if (lagoi>0) {
        flower_freq_df_standard_best<- flower_freq_df_standard_best %>% mutate(freq_sm=lag(freq_sm,n=lagoi)) %>% mutate(freq_sm=replace_na(freq_sm, 0))
      }  
      
      flower_freq_df_standard_best_list[[r]]<-flower_freq_df_standard_best %>% mutate(lag=lagoi)
    }
  }
  flower_freq_df_standard_best<-bind_rows(flower_freq_df_standard_best_list)
  
  # make plots
  flower_freq_comp<-ggplot(flower_freq_df_standard_best)+
    geom_point(aes(x=doy, y=npn, col="flower observation (USA-NPN)"), alpha=0.5)+
    geom_point(aes(x=doy, y=pollen, col="pollen count (NAB)"))+
    geom_line(aes(x=doy, y=pollen_clim, col="pollen count (NAB)"),alpha=0.5, lwd=1)+
    geom_point(aes(x=doy, y=evi, col="EVI (PS)"), alpha=0.2)+
    geom_line(aes(x=doy, y=freq_sm, col="flowering frequency"), lwd=1)+
    theme_classic()+
    facet_wrap(.~paste0(sitename, " (Lag: ", lag, ")")*year, ncol=4)+
    scale_color_manual(values=cols)+
    theme(legend.position="bottom")+
    theme(legend.title = element_blank())+
    ylab("")+
    ggtitle(paste0("Taxa: ",taxaoi," (Threshold: ",best_thres$direction, " ",best_thres$thres, ")"))
  flower_freq_comp
  jpeg(paste0(output_path, "flower frequency compared with other data.jpg"),
            height = 1600, width = 1600, res=150)
  print(flower_freq_comp)
  dev.off()
  
  flower_freq_corr<-ggplot(flower_freq_df_standard_best %>% mutate(year=as.factor(year)))+
    geom_point(aes(x=freq_sm, y=pollen , group=year, col=year))+
    geom_smooth(aes(x=freq_sm, y=pollen , group=year, col=year), method="lm", se=F, lwd=0.5)+
    geom_smooth(aes(x=freq_sm, y=pollen), method="lm")+
    theme_classic()+
    facet_wrap(.~sitename, ncol=4)+
    coord_equal()+
    xlim(0, 1)+
    ylim(0, 1)+
    ylab("pollen count^(1/2)")+
    xlab("flowering frequency")
  flower_freq_corr
  jpeg(paste0(output_path, "flower frequency and pollen count correlation.jpg"),
            height = 1200, width = 1600, res=150)
  print(flower_freq_corr)
  dev.off()
  
  across_site_and_year<-ggplot(flower_freq_df_standard_best %>%  
                                 mutate(year=as.factor(year)))+
    geom_point(aes(x=doy, y=pollen , group=year, col=year))+
    geom_line(aes(x=doy, y=freq_sm, group=year, col=year))+
    facet_wrap(.~paste0(sitename, " (Lag: ", lag, ")"), ncol=1)+
    theme_classic()+
    ylab("")+
    ggtitle(paste0("Taxa: ",taxaoi," (Threshold: ",best_thres$direction, " ",best_thres$thres, ")"))
  across_site_and_year
  
  jpeg(paste0(output_path, "phenology across site and year.jpg"),
            height = 1600, width = 1600, res=150)
  print(across_site_and_year)
  dev.off()
}
stopCluster(cl)
```

Display results in a Shiny app.
```{r, eval=TRUE, cache=FALSE}
shinyApp(

  ui = fluidPage(
    selectInput("taxa", "Taxa:", choices = taxa_list),
    selectInput("plot", "Plot:", choices = c("flower frequency compared with other data",
                                             "phenology across site and year",
                                             "flower frequency and pollen count correlation")),
    plotOutput("plot")
  ),

  server = function(input, output) {
    img_file<-reactive(paste0("/raid/users/ysong67/GitHub/phenology/RS4flower/output/data/",input$taxa,"/",input$plot, ".jpg"))
    output$plot = renderPlot({
      image_ggplot(image_read(img_file()))
    })
  },

  options = list(width = 1600, height=1600)
)
```

Compare accuracy with climatology.
```{r, eval=TRUE}
flower_freq_df_check_list<-vector(mode="list") 
for (taxaoi in taxa_list) {
  output_path<-paste0("/raid/users/ysong67/GitHub/phenology/RS4flower/output/data/", taxaoi,"/")
  flower_freq_df_check_list[[taxaoi]]<-read_rds(paste0(output_path,"accuracy check.rds")) %>% 
    mutate(taxa=taxaoi)
}
flower_freq_df_check<-bind_rows(flower_freq_df_check_list)
best_thres<-flower_freq_df_check %>% group_by(taxa, thres) %>% summarise(rmse=median(rmse)) %>% arrange(rmse) %>% slice(1) %>% dplyr::select(taxa, thres)

accuracy_df<-flower_freq_df_check %>% 
  right_join(best_thres, by=c("taxa", "thres")) %>% 
  left_join(meta_df %>% dplyr::select(site, sitename), by=c("region"="site"))

accuracy_df %>% 
  dplyr::select(taxa, site=region, sitename, direction, thres, lag)

ggplot(accuracy_df)+
  geom_point(aes(x=region, y=rmse_ps), col="dark blue")+
  geom_point(aes(x=region, y=rmse_clim), col="dark red")+
  facet_wrap(.~taxa)+
  ylab("RMSE")+
  theme_classic()

accuracy_df %>% 
  mutate(rmse_diff=rmse_clim-rmse_ps) %>% 
  drop_na() %>% 
  group_by(taxa) %>% 
  summarize (min=min(rmse_diff),
             max=max(rmse_diff),
             mean=mean(rmse_diff),
             median=median(rmse_diff),
             proportion=sum(rmse_diff>0)/n()) %>% 
  ungroup()
```

# 4 Infer relationship between leaf-flower lag with climate

Plot correlation between mean annual temperature (MAT) and lag between green-up/down frequency and pollen count.
* A positive lag means leafing phenology leads pollen phenology; a negative lag means leafing phenology lags pollen phenology.
```{r, eval=TRUE}
# taxa in chronological order
taxa_chron<-c("Cupressaceae" , "Fraxinus","Ulmus early"  ,"Pinaceae" ,"Acer", "Populus","Quercus",   "Betula","Morus",  "Poaceae early","Poaceae late"  ,    "Ulmus late" ,  "Ambrosia" )

# data frame with flowering frequency and climate info, grouped into early and late taxa
lag_clim_df<-accuracy_df %>% 
  dplyr::select(-rmse, -rmse_ps, -rmse_clim) %>% 
  left_join(chelsa_df, by=c("region"="site")) %>% 
  mutate(taxa=as.factor(taxa)) %>% 
  mutate(taxa=fct_relevel(taxa, levels=taxa_chron))%>%
  mutate(group=case_when(taxa %in% c("Ulmus late", "Poaceae late", "Ambrosia")~"late",
                                                                              TRUE~"early"))

# plot lag vs. climate
ggplot(lag_clim_df)+
  geom_point(aes(x=mat, y=lag, col=region, group=region))+
  geom_smooth(aes(x=mat, y=lag), method="lm", se=F)+
  geom_smooth(aes(x=mat, y=lag), method="lm", alpha=0.5)+
  theme_classic()+
  facet_wrap(.~taxa, scales = "free_y", ncol=5)+
  xlab("Mean annual temperature (°C)")+
  ylab ("Lag between leafing and pollen phenology (day)")
```
Linear regression to check significance of the correlation. When looking at individual taxa, only 3 were statistically significant.
```{r, eval=TRUE}
# linear regression
reg.df<-lag_clim_df  %>% 
  # filter(region!="SJ") %>% 
  group_by(taxa) %>%
  do(broom::tidy(lm(lag ~ mat, .))) %>%
  filter(term %in% c("mat")) %>%
  dplyr::select( -statistic)
ggplot(reg.df %>% mutate(taxa=fct_relevel(taxa,levels=rev(taxa_chron))))+
  geom_point(aes(x=taxa, y=estimate))+
  geom_errorbar(aes(x=taxa, ymin=estimate-1.95*std.error, ymax=estimate+1.95*std.error))+
  geom_hline(yintercept = 0, lty=2)+
  geom_vline(xintercept = 3.5, lty=1)+
  theme_classic()+
  coord_flip()+
  ylab("Slope (day/°C)")+
  xlab ("Taxa")
summary(reg.df$p.value<0.05)
```

Linear mixed-effect model with random slope and random intercept to summarize correlation over early and late flowering taxa.
* For early-flowering taxa, pollen phenology usually leads leafing phenology. At warmer places, pollen leads leafing even more. This effect was statistically significant.
* For early-flowering taxa, pollen phenology usually lags leafing phenology. At warmer places, pollen lags green-down even more. This effect was not statistically significant, possibly because there were only three taxa. It was significant for late-flowering elm trees.

```{r,eval=TRUE}
lme.fit<-lme(lag ~ mat, random = ~ 1+mat|taxa, data=lag_clim_df %>% filter(group=="early"),control =lmeControl(opt='optim',optimMethod = "SANN") )
summary(lme.fit)

lme.fit<-lme(lag ~ mat, random = ~ 1+mat|taxa, data=lag_clim_df %>% filter(group=="late"),control =lmeControl(opt='optim',optimMethod = "SANN"))
summary(lme.fit)
```

